### 2022-11-21 Draft_GFC.tex
21/11/2022

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL 
\usepackage{tabularx}
%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{-.3in}%
\addtolength{\topmargin}{-.8in}%
\usepackage{amsmath,amssymb,amsthm,amsfonts,amstext,amsbsy,amscd,stmaryrd}
\usepackage{amssymb,amsmath,comment,subcaption,graphicx}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{cancel}
\usepackage{tikz}
\usepackage{subcaption} 
\usetikzlibrary{patterns}
\usepackage{xcolor}
\usepackage{bbm}
\usepackage{float}
\usepackage{color}
\usepackage{relsize}
\newcommand{\modif}[1]{{\begin{modification}#1\end{modification}}}
\newcommand{\ele}[1]{{\color{red}#1}}
%\UseRawInputEncoding
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\usepackage{mathrsfs}
\newcommand{\rmd}{{\mathrm{d}}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\Z}{\mathbb{Z}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\operatorname{\mathbb{E}}}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\Cov}{\mathrm {Cov}}
\newcommand{\rme}{{\rm e}}
\renewcommand{\tilde}{\widetilde}
\renewcommand{\hat}{\widehat}
\newcommand{\TC}{\mathrm{TC}}
\newcommand{\TaC}{\mathrm{TaC}}
\newcommand{\LTC}{\mathrm{LTC}}
\newcommand{\LP}{\mathrm{LP}}
\newcommand{\Per}{\mathrm{Per}}
\theoremstyle{Theorem}
\newtheorem{Theorem}{Theorem}[section]
\newtheorem{Corollary}[Theorem]{Corollary}
\newtheorem{Lemma}[Theorem]{Lemma}
\newtheorem{Proposition}[Theorem]{Proposition}
\newtheorem{Definition}[Theorem]{Definition}
\newtheorem{remark}{Remark}
\renewcommand{\labelitemi}{\textperiodcentered}
\allowdisplaybreaks
\usepackage{ifpdf}
\ifpdf
\else
\usepackage{graphicx}
\DeclareGraphicsExtensions{.eps,.jpg,.png}
\fi
\usepackage{epstopdf}
\usepackage{array}
\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}

\newcommand{\overbar}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}
\newcolumntype{Y}{>{\centering\arraybackslash}X}

\parindent 0mm    
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}
\tikzset{
    cross/.pic = {
    \draw[rotate = 45] (-#1,0) -- (#1,0);
    \draw[rotate = 45] (0,-#1) -- (0, #1);
    }
}

\tikzset{
  mycross/.pic={
    \draw[pic actions] 
      (-2pt,0) -- (2pt,0)
      (0,-2pt) -- (0,2pt);
  },
}
\begin{document}
\section{Introduction}
\textbf{Cabana : Affine Processes: A test of isotropy based on level sets.} 
Class of affine processes for the alternative, $H_{0} :X$ is isotropic, create the test using Estimators for the affinity parameters based only on one level set of the observed process, show consistence under some hypothesis. Proposes estimators of the affinity parameters $k = (1 - \frac{\lambda_{1}^{2}}{\lambda_{1}^{2}})$ and $\theta_{0}$ using the shape of the level curve of $X$ for fixed level $x$. Gets TCL for estimators under condition of \textit{X is uniformly mixing}. \\
Let $\mathscr{C}$ is the level sets, principal objects that are used are $\mathscr{L}(T) = |T|^{-1}\int_{\mathscr{C}}ds$, $\mathscr{C}(T) = |T|^{-1}\int_{\mathcal{C}} \cos\Theta ds$, $\mathscr{S}(T) = |T|^{-1}\int_{\mathscr{C}} \sin2\Theta ds$ that are particular c*ases of line integrals : $\mathscr{F}(f, T) = |T|^{-1}\int_{\mathscr{C}}f(\Theta)ds$, with $f:(-\pi, \pi) \to \mathbb{R}$ is a bounded measurable function.\\
Weird assumptions imply that integrals such as $\mathscr{F}$ have a first and second moment. When we assume that the field $X(t) = Y(At)$ with $Y$ isotropic and we substitute $f$ with the function that are of interest for us, we get a formula for the first moment of $\mathscr{L}$, $\mathscr{C}$ and $\mathscr{S}$. that only depends on $\lambda_{1}, k$ and $\theta_{0}$ and a weird function $g$ that is a continuously increasing function. Thus, we get that $\tan 2\theta_{0} = E\mathscr{S}/E\mathscr{C}$ and $g(k) = \sqrt{(E\mathscr{C})^2 + (E\mathscr{S})^2}/E\mathscr{L}$, \textbf{Asymptotic distributions}. Assuming the field $X$ is uniformly mixing, then we have a TCL for $\mathscr{F}(f, T)$ when the window goes to infinity, that give them the fact that $\hat{\theta_{0}}$ and $\hat{k}$ are consistent. $k$ measures the departure from isotropy. I didn't understand the last paragraph of page 5. But something leads them to consider an $F$ test of isotropy. Considering a horrible test variable $\mathscr{B}(\mathscr{T})$ with $\mathscr{T}$ is partition of $T$, they know that its distribution under $H_{0}$ ($k = 0$) tends to $F_{2, 2n-2}$ $n$ the number of partitions, if $X$ is an affine process, but not isotropic ($k \neq 0$) than the limit law is a non central F, as it is apparently well known from the analysis of variance. The critical region $\mathscr{B}(\lambda \mathscr{T}) > F_{2, 2n-2}(1-\alpha)$ ($\lambda \to \infty$ is an expansion parameter). \\
\textbf{Molina and al. : A method for testing anisotropy and quantifying its direction in digital images.}\\
 \textit{The identification of appropriate spatial models require the previous knowledge of the process exploratory properties such as the existence of a directional component which determines an anisotropic behavior of the phenomenon. Techniques used for analyzing anistropy are limited (compared to techniques for exploring stationarity) and they are only devoted to examining "by eye" experimental variograms in
different directions. (I do not know if that is true or not).} \\
They try to detect anisothropy using a method based on second order bi variate \textit{circular statistics} this allows them also to quantify the direction in which it appears. They also present an application of the method to  the flow of seawater through the Strait of Gibraltar. Tools : a parametric procedure based on the calculation of standard and confidence ellipses, well how do they do that ? and what is that. Well, It's pretty weird, they start by considering a sub image $Z_{l,l}$ of the image $Z$ and $Z(s)$ the random variable (yep nothing more on this mystery variable ), from $Z_{l,l}$ they draw $N=\{z_{1}, \ldots, z_{m}\}$ a random sample of $m$ elements, nothing is being said about this random sample. $\overrightarrow{z_{i}z_{j}}$ represents a variation of $Z(s)$ with intensity $|z_{j} - z_{i}|$ in the direction $\gamma$, being 
\begin{equation*}
\gamma = \begin{cases} \arctan{\frac{z_{jy} - z_{iy}}{z_{jx} - z_{ix}}}& \text{if} (z_{j} - z_{i}) > 0, \\
0 & \text{if} (z_{j} - z_{i}) = 0, \\
 \arctan{\frac{z_{jy} - z_{iy}}{z_{jx} - z_{ix}}} + \pi & \text{if} (z_{j} - z_{i})  < 0, 
\end{cases}
\end{equation*}
where $(z_{ix}, z_{iy})$ are the coordinates of intensity $z_i$, same for the other one. Taking the vectors $\overrightarrow{z_{i}z_{j}}$ we obtain the set $\Gamma = \{\gamma_{i}, i = 1, \ldots, n, n = \binom{m}{2}\}$, that contains the orientations in the plane of vectors. Then, they proceed to the computation of the mean vector $\hat{m}$. Given a sample $\Gamma$ its mean vector $\hat{m}$ (of length $r$ and mean angle $\hat{\Phi}$) is calculated as follows : 
\begin{equation*}
\hat{x} = \frac{1}{n} \sum_{i}^{n} b_{i}\cos{\gamma_{i}}, \; \hspace{3cm}  \; \hat{y} = \frac{1}{n}\sum_{i}^{n}b_{i}\sin{\gamma_{i}}
\end{equation*}
\begin{equation*}
r=\sqrt{\hat{x}^{2} + \hat{y}^{2}}, \; \hspace{1cm} \; \hat{\Phi} = \begin{cases} \arctan{\frac{\hat{y}}{\hat{x}}} \hspace{0.5cm} \text{if} \; \hat{x} > 0 , \\
180 + \arctan{\frac{\hat{y}}{\hat{x}}} \hspace{0.5cm} \text{if} \; \hat{x} < 0 
\end{cases}
\end{equation*}We debi
By using standard ellipse, they can compute the direction of maximum variability in a direction of angle $\theta$. So, they are able to describe the spatial behavior of the mean vectors, quantifying by the calculation of $\theta$ the direction of maximum variability

\section{Mathematical framework}\label{section1}
\subsection{Construction of the binary image}\label{construction}
\paragraph{Square tiling}
Let $m$ be an integer with $m \geq 2$. Without loss of generality, we consider our observation window as the unit square $S = [0,m]^{2}$ and  we divide our window into $m^{2}$ pairwise disjoint squares. We denote by $\mathbb{G}_{m} := \text{\larger[1]{$[$}}0, m \text{\larger[1]{$)$}}^{2} \cap \mathbb{Z}^{2} $ the set of point in $S$ with interger coordinates and for $x \in \mathbb{G}_{m}$, we define $C_{m}(x)$ as
\begin{equation*}
C_{m}(x) \, := \,x + \text{\larger[1]{$[$}}0, 1 \text{\larger[1]{$]$}}^{2} \subset S.
\end{equation*}
The $C_{m}(x)$ will be referred to as \textit{cells}, to ease up the notations we denote by $x = (x_1, x_2)$ with $x_1, x_2 \in \{0,\ldots, m-1\}$ and $e_1$, $e_{2}$ the element of the canonical base of $\mathbb{R}^{2}$.
We denote by $\mathbb{E}^{(1)}_{m}, \; \mathbb{E}^{(2)}_{m} $ the set of vertical and horizontal edges in $\mathring{S} = (0,m)^{2}$, and $\mathbb{E}_{m} = \mathbb{E}^{(1)}_{m} \cup \mathbb{E}^{(2)}_{m}$ each $w \in \mathbb{E}_{m}$ is a segment of length $1$.
Figure \ref{fig1} shows an example of a square tiling with $m = 4$.
\begin{figure}[h]
\begin{center}
\begin{tikzpicture}[scale = 0.5]
  \draw[step = 1cm, black, thick](0,0) grid (4,4);
\draw[pattern=north west lines, pattern color=blue] (0,0) rectangle (1,1);
\draw[pattern=north west lines, pattern color=blue] (3,1) rectangle (2,2);
\foreach \x in {0,1,2,3}{
  \foreach \y in {0,1,2,3}{
  \draw[fill = red] (\x,\y) circle (2pt);
  } 
}
\end{tikzpicture}
\end{center}
\vspace{-0.25cm}
\caption{Tiling with squares for $m = 4$ and associated cells $C_{4}(0,0)$ and $C_{4}(2,1)$ (blue stripes) and red point of $\mathbb{G}_{m}$.}
\label{fig1}
\end{figure}

\paragraph{Discrete setting}
Using the previous square tiling, we write $S =  \underset{x \in \mathbb{G}_{m}}{\cup}  C_{m}(x)$. We assume to observe $\left(X_{\scriptscriptstyle x}\right)_{x \in \mathbb{G}_{m}}$, which we assume coincides with values of a stationary random field $\left(X_{\scriptscriptstyle x}\right)_{x \in \mathbb{Z}^{2}}$. The stationarity hypothesis means that for all $h \in \mathbb{Z}^{2}$, 
$$\left(X_{x+h}; x \in \mathbb{Z}^{2} \right)\overset{fdd}{=} \left(X_{\scriptscriptstyle x}; x \in \mathbb{Z}^{2} \right),$$ where $fdd$ denotes finite-dimensional distributions. \\
We denote by $\rho(x) :=\text{Cov}\left(X_{\scriptscriptstyle 0}, X_{\scriptscriptstyle x}\right), \; x \in \mathbb{Z}^{2}$, the covariance function of $X$.
\vspace{-1.5cm}
\begin{figure}[H]
    {\includegraphics[scale=0.25]{GaussianUn.pdf}}
    {\includegraphics[scale=0.25]{GaussianZUn.pdf}}
    {\includegraphics[scale=0.25]{GaussianZZZUn.pdf}}
    \vspace{-2cm}
 \caption{Generations of Gaussian random field with covariance $\rho(x) = e^{-\kappa||x||^{2}}$ with $m = 1024$, and different values for $\kappa$, from left to right $\kappa = 0.01, 0.1, 1$, using the Matlab function \textit{stationary Gaussian process} \cite{MATLAB}.}
\label{fig22}
\end{figure}
Let us consider a threshold parameter $t \in \mathbb{R}$. We introduce the associated binary image $Z^{\scriptscriptstyle (m)}_{t} = \left(\mathbbm{1}_{\left\{X_x \geq t\right\}} \right)_{x \in \mathbb{G}_{m}}$, each cell $C_{m}(x)$ is associated to black or white according to whether $Z^{\scriptscriptstyle (m)}_{t} (x) = 0$ or $Z^{\scriptscriptstyle (m)}_{t}(x)~= 1$. 
\subsection{Perimeter of a binary image}\label{methods}
For $t \in \mathbb{R}$,let $Z^{\scriptscriptstyle (m)} = Z^{\scriptscriptstyle (m)}_{t}$ be the  binary image  $Z^{\scriptscriptstyle (m)}$ at that given threshold $t \in \mathbb{R}$. Following  the approach  presented in \cite{HermineAgnes},  for each edge $w \in \mathbb{E}_{m}$, we aim to know if  $w$ contributes  to the perimeter of the black component of $Z$. Making use of the additive nature of the perimeter, one can start by computing the oriented perimeter given each direction, horizontal and vertical. Let us consider the edge $w_{i}(x) := \left(x, x+e_{i}\right)$ with $i \in \{1,2\}$, $w_{i}(x)$ belongs to $C_{m}(x) \cap C_{m}(x-e_{j})$ with $i,j \in \{1,2\}$ and $i \neq j$ as soon as $w_{i}(x) \subset  \text{\larger[1]{$($}}0, m \text{\larger[1]{$)$}}^{2}$.
\begin{figure}[H]
\begin{center}
\begin{tikzpicture}
    [
        box/.style={rectangle,draw=black,thick, minimum size=1cm},
    ]

    here the test 
\foreach \x in {0,1}{
    \foreach \y in {0,1}
        \node[box] at (\x,\y){};
}

\node[box, fill=black] at (0,0){};  
\node[box, pattern = north west lines, pattern color=gray] at (1,0){};  
\draw[step = 1cm, red, thick] (0.5,-0.5) -- (0.5,0.5);
\foreach \x in {0.5}
{
  \draw 
    (\x cm, 0pt) -- (\x cm, 0pt) node[anchor=north] {\scriptsize $\text{\textcolor{red}{}}$};
  \pic[line width=0.5pt,red, rotate = 45] at (\x, 0.5) {mycross};
  \pic[line width=0.5pt,red, rotate = 45] at (\x, -0.5) {mycross};  
}
\end{tikzpicture}
\end{center}
\vspace{-0.25cm}
\caption{The edge $w_{1} = (1,1)$ in red belongs to $C_{2}(0,0) \cap C_{2}(1, 0)$, with the cell $C_{2}(0,0)$ colored in black and $C_{2}(1, 0)$ in gray dashed stripes.}
\label{fig1}
\end{figure}
\vspace{-0.5cm}
Following this consideration, one can define the random quantity
\begin{equation}
f_{t}^{\scriptscriptstyle (i)}(x) := \mathbbm{1}_{\left\{\min\left(X_{x}, X_{\scriptscriptstyle x +e_{i}}\right) < t \leq \max\left(X_{x}, X_{\scriptscriptstyle x+e_{i}}\right)\right\}}
\label{fi}
\end{equation}
to compute the contribution of the edge $w_i(x)$ to the perimeter of the black component of $Z^{\scriptscriptstyle (m)}$.
\begin{Definition}[Perimeter of a binary image $Z^{\scriptscriptstyle (m)}$]\label{defPerimetre}~\\
For $t\in \mathbb{R},$ we denote by $\mathcal{P}_{m}^{\scriptscriptstyle (i)}(t) =~\sum_{\scriptscriptstyle x: w_{i}(x) \in \mathbb{E}^{\scriptscriptstyle (i)}_{m}} f_{t}^{\scriptscriptstyle (i)}(x)$ the  sum of all vertical contributions for $i = 1$ and horizontal ones for $i = 2$. Then, the perimeter of the binary image $Z$ is given by
\begin{equation}
\mathcal{P}_{m}(t) := \mathcal{P}_{m}^{\scriptscriptstyle (1)}(t) + \mathcal{P}_{m}^{\scriptscriptstyle (2)}(t).
 \label{eq:6}
\end{equation}
\end{Definition}
\section{Statistics of the perimeter} 
\subsection{First moment}
Let us start this study by investigating the first moment of the perimeter.
\begin{Proposition}[First moment of the perimeter] 
\label{propFisrtmoment}
We assume that $\left(X_x \right)_{x \in \mathbb{G}_{m}}$ is a stationary Gaussian centered with unit variance random field let us denote \linebreak by  $\rho(e_i) = \text{Cov}\left(X_{\scriptscriptstyle 0}, X_{e_i}\right) \in~[-1, 1]$. The expected value of the perimeter in \eqref{eq:6} is  given \linebreak by $$\mathbb{E}\left(\mathcal{P}_{m}(t)\right) = m(m-1)\sum_{1}^{2}\mathbb{E}\left(f^{(i)}(t) \right)$$  and
\begin{itemize}
  \item if $\rho(e_{i}) \in ]-1,1]$, 
    {\small
  \begin{align}
  \label{propEquationFisrtmoment}
  \mathbb{E}\left(f^{(i)}(t) \right) = \mathbb{E}\left(\Phi\left(\dfrac{2t + \sqrt{2(1-\rho(e_i))}|N|}{\sqrt{2(1+\rho(e_i))}}\right)  - \Phi\left(\dfrac{2t - \sqrt{2(1-\rho(e_i))}|N|}{\sqrt{2(1+\rho(e_i))}}\right)\right)
  \end{align}}
with $N \sim \mathcal{N}(0,1)$. 
\item if $\rho(e_i) = -1, \; \mathbb{E}\left(f^{(i)}(t) \right)= 2\left(1 - \Phi\left(\max\left(0,t\right)\right)\right).$
  \end{itemize}
\end{Proposition}
\begin{remark}
Note that $\mathbb{E}\left(f^{(i)}(t) \right) = 0$ in the case where $\rho(e_i) = 1$ 
\end{remark}
\begin{proof}
Let us assume that $\rho(e_{i}) \in ]-1,1]$. Considering the formula for $\mathcal{P}^{\scriptscriptstyle  (i)}_{m}(t)$ and applying the stationarity hypothesis, since
\begin{align*}
\left\{ x \in \mathbb{G}_{m}; w_{i}(x) \in \mathbb{E}^{\scriptscriptstyle  (i)}_{m}\right\} & = \left\{x \in \llbracket 0, m-1 \rrbracket^{2}, (x, x+e_i)\subset (0,m)^{2} \right\} \\ 
& = \left\{ 0, m-1 \right\} \times \left\{ 1, m - 1\right\},
\end{align*}
one can write that, 
\begin{equation*}
\mathbb{E}\left(\mathcal{P}^{\scriptscriptstyle  (i)}_{m}(t) \right) = m(m-1)\mathbb{E}\left(f^{(i)}(t)\right).
\label{covarianeproof}
\end{equation*}
 with $\mathbb{E}\left(f^{(i)}(t)\right) = \mathbb{P}\left(\min\left(X_{\scriptscriptstyle 0}, X_{e_i}\right) < t \leq \max\left(X_{\scriptscriptstyle 0}, X_{e_i}\right)\right)$. \\
 We start by applying to the Gaussian vector $\left(X_{\scriptscriptstyle 0}, X_{e_i}\right)$ the following transformation \begin{equation*} \label{expressions} \Delta^{\scriptscriptstyle (i)}_{\scriptscriptstyle 0} := X_{e_i} - X_{\scriptscriptstyle 0}, \hspace{0.5cm} S^{\scriptscriptstyle (i)}_{\scriptscriptstyle 0} := X_{e_i} + X_{\scriptscriptstyle 0}\end{equation*}
Let us introduce the following Lemma, 
\begin{Lemma}
\label{lemmasommediff}
Let $a,b,t \in \mathbb{R}$, then 
$$\min(a,b) \leq t \leq \max(a,b) \iff\left|t - \frac{a+b}{2}\right| \leq \left| \frac{ b-a }{2}\right| $$
\end{Lemma}
\begin{proof}
Let $a,b,t \in \mathbb{R}$,
\begin{figure}[H]
\centering
\begin{tikzpicture}[scale = 1.5]
\draw (-1,0)--(9,0);
\draw (0,-0.08)--(0,0.08);
\draw (8,-0.08)--(8,0.08);
\draw (2.5, -0.08) -- (2.5, 0.08);
\draw (4, -0.08) -- (4, 0.08);
\draw (0,0) node[below]{$\min(a,b)$};
\draw (8,0) node[below]{$\max(a,b)$};
\draw (4,0) node[below]{$\frac{b+a}{2}$};
\draw (2.5,0) node[below]{$t$};
\draw (2.5,-0.6) -- (4,-0.6);
\draw (2.5, -0.6) -- (2.75,-0.5);
\draw (2.5, -0.6) -- (2.75,-0.7);
\draw (4, -0.6) -- (3.75,-0.5);
\draw (4, -0.6) -- (3.75,-0.7);
\draw (0,0.6) -- (8,0.6);
\draw (0, 0.6) -- (0.25,0.5);
\draw (0, 0.6) -- (0.25,0.7);
\draw (8, 0.6) -- (7.75,0.7);
\draw (8, 0.6) -- (7.75,0.5);
\node[text width= 1.6cm, font=\large] at (4,1) {$\frac{|b-a|}{2}$};
\node[text width= 0.7cm, font=\large] at (3,-1) {$|t -~\frac{b+a}{2}|$};
\end{tikzpicture}
\end{figure}

\end{proof}
Thus, applying the Lemma we get 
{\small
\begin{align*}
\mathbb{E}\left(f^{(i)}(t)\right) & = \mathbb{P}\left(\min\left(X_{\scriptscriptstyle 0}, X_{e_i}\right) < t \leq \max\left(X_{\scriptscriptstyle 0}, X_{e_i}\right)\right) \\
& = \mathbb{E}\left(\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle 0}, X_{e_i}) \leq t \leq \max(X_{\scriptscriptstyle 0}, X_{e_i})\right\}}\right) \\
& = \mathbb{E}\left(\mathbbm{1}_{\left\{ \left|2t - S^{\scriptscriptstyle (i)}_{\scriptscriptstyle 0}\right| \leq \left|\Delta^{\scriptscriptstyle (i)}_{\scriptscriptstyle 0}\right|\right\}}\right).
\end{align*}}Besides, the covariance matrix of the new Gaussian vector $\left(\Delta^{\scriptscriptstyle (i)}_{\scriptscriptstyle 0}, S^{\scriptscriptstyle (i)}_{\scriptscriptstyle 0}\right)$ is given by ${\tilde{\Sigma}_{i}(0)~= M\Sigma_{i}(0) M^{\star}}$ with {\small $M = \begin{pmatrix}
-1 & 1 \\
1 & 1  \\
\end{pmatrix}$} and $\Sigma_{i}(0) = \begin{pmatrix} 1 & \rho(e_i) \\
\rho(e_i) & 1  \\ 
\end{pmatrix}.$ \\
Thus, $\tilde{\Sigma}_{i}(0) = \begin{pmatrix}  2(1-\rho(e_i)) & 0 \\ 0 & 2(1+\rho(e_i)  \end{pmatrix}$, which directly  implies that the two variables~$\Delta^{\scriptscriptstyle (i)}_{\scriptscriptstyle 0}$ and $S^{\scriptscriptstyle (i)}_{\scriptscriptstyle 0}$ are independent, we get the following system 
$\left\{
 \begin{array}{rlc}
  \Delta^{\scriptscriptstyle (i)}_{\scriptscriptstyle 0} & = & \sqrt{2(1-\rho(e_i))}U, \textcolor{white}{(1-\rho(e_i))U + }\\
  S^{\scriptscriptstyle (i)}_{\scriptscriptstyle 0} & = & \textcolor{white}{(1-\rho(e_i))U + } \sqrt{2(1+\rho(e_i)}V, 
\end{array}
\right.$
{\small
\begin{align*}
\mathbb{E}\left(f^{(i)}(t)\right) & = \mathbb{E}\left(\mathbbm{1}_{\left\{ \left|2t - S^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}\right| \leq \left|\Delta^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}\right|\right\}}\right) \\
& = \mathbb{E}\left(\mathbbm{1}_{\left\{ - \sqrt{2(1-\rho(e_i))}\left|U\right| \leq 2t - \sqrt{2(1+\rho(e_i)}V \leq \sqrt{2(1-\rho(e_i))}\left|U\right| \right\}}\right) \\
& = \mathbb{E}\left(\mathbbm{1}_{\left\{\frac{2t - \sqrt{2(1-\rho(e_i))}|U|}{\sqrt{2(1+\rho(e_i))}}  \leq V \leq \frac{2t + \sqrt{2(1-\rho(e_i))}|U|}{\sqrt{2(1+\rho(e_i))}}\right\}}\right), 
\end{align*}}  
with $U, V \sim \mathcal{N}(0,I_{2})$, thus 
$$\mathbb{E}\left(f^{(i)}(t)\right) = \mathbb{E}\left(\Phi\left(\dfrac{2t + \sqrt{2(1-\rho(e_i))}|N|}{\sqrt{2(1+\rho(e_i))}}\right)  - \Phi\left(\dfrac{2t - \sqrt{2(1-\rho(e_i))}|N|}{\sqrt{2(1+\rho(e_i))}}\right)\right).$$
Let us now consider the case where $\rho(e_i) = -1$. Then there exists $N \sim \mathcal{N}\left(0,1\right)$ such that $\left(X_{\scriptscriptstyle 0}, X_{e_i}\right) \overset{d}{=} \left(N, -N\right)$. Thus, 
\begin{align*}
\mathbb{P}\left(\min\left(X_{\scriptscriptstyle 0}, X_{e_i}\right) < t \leq \max\left(X_{\scriptscriptstyle 0}, X_{e_i}\right)\right) & = \mathbb{P}\left(\min\left(N, -N\right) < t \leq \max\left(N, -N\right)\right) \\
& = \mathbb{P}\left(|t| \leq |N|\right) = 2\left(1 - \Phi\left(|t|\right)\right).
\end{align*}
\end{proof}
Note that this result extends those given in \cite{HermineAgnes} (Proposition 3) since we do not assume here that $\rho(e_i) > 0$. 
\begin{remark}
\label{iidCase}
In the independent case, $\rho(e_i) = 0$, we can write
{\small
\begin{align*}
\mathbb{E}\left(\mathcal{P}^{\scriptscriptstyle  (i)}_{m}(t) \right) & = m(m-1)\mathbb{E}\left(\Phi\left(\sqrt{2}t + |Z|\right) - \Phi\left(\sqrt{2}t - |Z|\right)\right) = 2m(m-1)\Phi(t)\left(1-\Phi(t)\right) 
\end{align*}}
by applying a variable change {\small$\left\{\begin{array}{rcl}
u & = & \frac{y - z}{\sqrt{2}} \\
v & = & \frac{y+z}{\sqrt{2}} \\
\end{array}\right.$}, we get the same result as in \cite{Psymetrie} (Proposition \textcolor{blue}{3.1}).
\end{remark}
\begin{remark}
In order to have convergence when $m$ goes to infinity, one has to normalize the perimeter with the area of the window ($m^2$), considering $\frac{1}{m^2}\mathcal{P}_{m}(t)$.
\end{remark}
The total variation of an image $u \in L^{1}_{loc}(\Omega)$ is given by $\int_{\Omega}|\nabla u|dx$ which is equal by the coarea formula to $\int_{\mathbb{R}}\mathcal{P}(t)dt$ with $\mathcal{P}(t)$ being the perimeter associated to the level set of the image $u$ at level $t$ .
\begin{Proposition}
Under the assumptions stated in Proposition \ref{propFisrtmoment}, we get 
\begin{equation}
\label{totalvariation}
\int_{\mathbb{R}} \mathbb{E}\left(\mathcal{P}^{(i)}_{m}(t)\right) dt = \frac{2m(m-1)}{\sqrt{\pi}}\left(\sqrt{1-\rho(e_i)}  \right).
\end{equation}
\end{Proposition}
\begin{proof}
Using the result stated in Proposition \ref{propFisrtmoment}, we get that $\mathbb{E}\left(\mathcal{P}^{\scriptscriptstyle  (i)}_{m}(t) \right)$ is equal to,
{\small
\begin{align*}
\int_{\mathbb{R}} \mathbb{E}\left(\mathcal{P}^{(i)}_{m}(t)\right) dt & = \\
& \hspace{-2cm}\cfrac{m(m-1)}{\sqrt{2\pi}} \int_{\mathbb{R}}\int_{\mathbb{R}} \left(\Phi\left(\dfrac{\sqrt{2}t + \sqrt{1-\rho(e_i)}|u|}{\sqrt{1+\rho(e_i)}}\right) - \Phi\left(\dfrac{\sqrt{2}t - \sqrt{1-\rho(e_i)}|u|}{\sqrt{1+\rho(e_i)}}\right)\right)e^{-\frac{1}{2}u^2}dudt \\
& \hspace{-2cm}=\cfrac{m(m-1)}{2\pi} \int_{\mathbb{R}}\int_{\mathbb{R}}\int_{\mathbb{R}} e^{-\frac{1}{2}u^2}e^{-\frac{1}{2}v^2} \mathbbm{1}_{\left\{\dfrac{\sqrt{2}t - \sqrt{1-\rho(e_i)}|u|}{\sqrt{1+\rho(e_i)}}\leq v \leq\dfrac{\sqrt{2}t + \sqrt{1-\rho(e_i)}|u|}{\sqrt{1+\rho(e_i)}}\right\}} du dv dt ,\\
& \hspace{-2cm}\text{using Fubini-Tonelli Theorem,} \\
& \hspace{-2cm}= \cfrac{m(m-1)}{2\pi}\int_{\mathbb{R}}\int_{\mathbb{R}} e^{-\frac{1}{2}u^2}e^{-\frac{1}{2}v^2} \\
& \hspace{-2cm} \left(\int_{\mathbb{R}} \mathbbm{1}_{\left\{\dfrac{\sqrt{1+\rho(e_i)}v - \sqrt{1-\rho(e_i)}|u|}{\sqrt{2}}\leq t\leq\dfrac{\sqrt{1+\rho(e_i)}v + \sqrt{1-\rho(e_i)}|u|}{\sqrt{2}}\right\}} dt\right) du dv \\
& \hspace{-2cm}= \cfrac{m(m-1)}{2\pi} \int_{\mathbb{R}}\int_{\mathbb{R}} e^{-\frac{1}{2}u^2}e^{-\frac{1}{2}v^2}\left(\dfrac{2|u|\sqrt{1-\rho(e_i)}}{\sqrt{2}}\right)dudv = \frac{2\sqrt{1-\rho(e_i)}}{\sqrt{2}} m(m-1)\underbrace{\mathbb{E}(|U|)}_{\sqrt{\frac{2}{\pi}}}\\
&\hspace{-2cm}=\frac{2}{\sqrt{\pi}}m(m-1)\sqrt{1-\rho(e_i)}.
\end{align*}
}
\end{proof}
\begin{remark}
Using this result, one can define an estimator for $\rho(e_i)$. 
\end{remark}
\begin{remark}
\label{decreasingrho}
Considering $\tilde{\rho}(e_i), \bar{\rho}(e_i)$ two covariance, we  have the following result concerning the expected perimeter of the nodal curve ($t=0$).
\begin{equation*}
 \text{if} \hspace{0.5cm} \tilde{\rho}(e_i) < \bar{\rho}(e_i) \hspace{0.5cm} \text{then} \hspace{0.5cm}\mathbb{E}\left(\mathcal{P}^{\scriptscriptstyle (i)}_{m, \tilde{\rho}}(0)\right) > \mathbb{E}\left(\mathcal{P}^{\scriptscriptstyle (i)}_{m, \bar{\rho}}(0)\right),
\end{equation*}
where $\mathcal{P}^{\scriptscriptstyle (i)}_{m, \tilde{\rho}}$ is the perimeter with respect of the orientation $i$ associated to the Gaussian field of covariance $\tilde{\rho}$. 
\end{remark}
\begin{proof}
One has $\mathbb{E}\left(f^{(i)}(0)\right) = \mathbb{E}\left(\frac{\sqrt{1-\rho(e_i)}}{\sqrt{1+\rho(e_i)}}|N|\right) - \mathbb{E}\left(-\frac{\sqrt{1-\rho(e_i)}}{\sqrt{1+\rho(e_i)}}|N|\right) = 2\mathbb{E}\left(\frac{\sqrt{1-\rho(e_i)}}{\sqrt{1+\rho(e_i)}}|N|\right) -~1$. Since $\rho \mapsto \frac{\sqrt{1-\rho}}{\sqrt{1+\rho}}$ is strictly decreasing, we see that $\rho \mapsto 2\mathbb{E}\left(\frac{\sqrt{1-\rho}}{\sqrt{1+\rho}}|N|\right) - 1 $ is also strictly decreasing. 
\end{proof}
\begin{remark}
Considering $\tilde{\rho}(e_i), \bar{\rho}(e_i)$ two covariance, we  have the following result concerning the expected perimeter at level $t$:
\begin{equation*}
 \text{if} \hspace{0.5cm} \tilde{\rho}(e_i) \leq \bar{\rho}(e_i) \hspace{0.5cm} \text{then} \hspace{0.5cm}\mathbb{E}\left(\mathcal{P}^{\scriptscriptstyle (i)}_{m, \tilde{\rho}}(t)\right) \geq \mathbb{E}\left(\mathcal{P}^{\scriptscriptstyle (i)}_{m, \bar{\rho}}(t)\right).
\end{equation*}
Where $\mathcal{P}^{\scriptscriptstyle (i)}_{m, \tilde{\rho}}$ is the perimeter with respect of the orientation $i$ associated to the Gaussian field of covariance $\tilde{\rho}$. 
\end{remark}
\begin{proof}
Let $t \in \mathbb{R}$ and $X$, $Y$ two centered with unit variance Gaussian random fields on $\mathbb{Z}^{2}$ of covariance stucture respectively $\tilde{\rho}$ and $\bar{\rho}$. We assume that $\tilde{\rho}(e_i) \leq \bar{\rho}(e_i)$, and consider the Gaussian vectors $\left(X_{0}, X_{e_1}\right)$ and $\left(Y_{0}, Y_{e_1}\right)$, $\tilde{\rho}(e_i) \leq \bar{\rho}(e_i)$ implies that $\mathbb{E}\left(X_{0}X_{e_1}\right) < \mathbb{E}\left(Y_{0}Y_{e_1}\right)$, we can write that
\begin{align*}
\mathbb{P}\left(Y_{\scriptscriptstyle 0} < t \leq  Y_{e_1}\right) = \mathbb{P}\left(Y_{\scriptscriptstyle 0} < t\right) - \mathbb{P}\left(Y_{\scriptscriptstyle 0} < t,  Y_{e_1} < t\right),
\end{align*}
according to Slepian's lemma, $\mathbb{P}\left(X_{\scriptscriptstyle 0} < t,  X_{e_1} < t\right) \leq \mathbb{P}\left(Y_{\scriptscriptstyle 0} < t,  Y_{e_1} < t\right)$, \linebreak besides,~$\mathbb{P}\left(Y_{\scriptscriptstyle 0} < t\right)=\mathbb{P}\left(X_{\scriptscriptstyle 0} < t\right)$ thus, 
\begin{align*}
\mathbb{P}\left(Y_{\scriptscriptstyle 0} < t \leq  Y_{e_1}\right) \leq \mathbb{P}\left(X_{\scriptscriptstyle 0} < t \leq  X_{e_1}\right)
\end{align*}
and,
\begin{align*}
\mathbb{P}\left(\min\left(Y_{\scriptscriptstyle 0}, Y_{e_1}\right) < t \leq \max\left(Y_{\scriptscriptstyle 0}, Y_{e_1}\right)\right)  \leq \mathbb{P}\left(\min\left(X_{\scriptscriptstyle 0}, X_{e_1}\right) < t \leq \max\left(X_{\scriptscriptstyle 0}, X_{e_1}\right)\right) 
\end{align*}
thus, 
\begin{align*}
\mathbb{E}\left(\mathcal{P}^{\scriptscriptstyle (i)}_{m, \tilde{\rho}}(t)\right) \geq \mathbb{E}\left(\mathcal{P}^{\scriptscriptstyle (i)}_{m, \bar{\rho}}(t)\right).
\end{align*}
\end{proof}
\begin{remark}
Considering the contraposition of this result, we get the following 
\begin{equation*}
 \text{if} \hspace{0.5cm}\mathbb{E}\left(\mathcal{P}^{\scriptscriptstyle (i)}_{m, \tilde{\rho}}(t)\right) \neq \mathbb{E}\left(\mathcal{P}^{\scriptscriptstyle (i)}_{m, \bar{\rho}}(t)\right)   \hspace{0.5cm} \text{then}  \hspace{0.5cm} \tilde{\rho}(e_i) \neq \bar{\rho}(e_i) \hspace{0.5cm}.
\end{equation*} 
\end{remark}
\subsection{Covariance of the perimeter for two thresholds}
Let $s, t \in \mathbb{R}$ such that $t \leq s$. Let us consider a stationary Gaussian random field $\left(X_{\scriptscriptstyle x}\right)_{x \in \mathbb{Z}^{2}}$ with zero mean and unit variance whose covariance structure is given by $\rho(x) = \text{Cov}\left(X_{\scriptscriptstyle 0}, X_{\scriptscriptstyle x}\right)$.
\begin{Lemma} 
\label{CovLemma}
\begin{align}
\text{Cov}(\mathcal{P}_{m}(t), \mathcal{P}_{m}(s))  & = \text{Cov}(\mathcal{P}_{m}^{\scriptscriptstyle (1)}(t), \mathcal{P}_{m}^{\scriptscriptstyle (1)}(s)) + \text{Cov}(\mathcal{P}_{m}^{\scriptscriptstyle (1)}(t), \mathcal{P}_{m}^{\scriptscriptstyle (2)}(s)) + \text{Cov}(\mathcal{P}_{m}^{\scriptscriptstyle (2)}(t), \mathcal{P}_{m}^{\scriptscriptstyle (1)}(s)) \\
& + \text{Cov}(\mathcal{P}_{m}^{\scriptscriptstyle (2)}(t), \mathcal{P}_{m}^{\scriptscriptstyle (2)}(s)) .
\label{covariance}
\end{align} 
With 
\begin{align*}
\text{Cov}\left(\mathcal{P}_{m}^{\scriptscriptstyle (i)}(t),\mathcal{P}_{m}^{\scriptscriptstyle (i)}(s) \right) 
& = \sum_{x_{\scriptscriptstyle j}=\left(1-m\right)}^{m-1}\sum_{x_{\scriptscriptstyle i}=\left(2-m\right)}^{m-2}\left(m -|x_{j}|\right)\left(m - 1- |x_{i}|\right) \text{Cov}\left(f_{t}^{\scriptscriptstyle (i)}(0), f_{s}^{\scriptscriptstyle (i)}(x) \right) 
\end{align*} 
for $i,j \in\{1,2\}, i\neq j$, and 
\begin{align*}
\text{Cov}\left(\mathcal{P}_{m}^{\scriptscriptstyle (1)}(t),\mathcal{P}_{m}^{\scriptscriptstyle (2)}(s) \right) 
& = \sum_{x_1=2-m}^{m-1} \sum_{x_2 = 1-m}^{m-2} \left(m-|x_1| - \mathbbm{1}_{x_{1} \leq 0}\right)\left(m-|x_2|-\mathbbm{1}_{x_{2} \geq 0}\right)\text{Cov}\left(f_{t}^{\scriptscriptstyle (1)}(0), f_{s}^{ \scriptscriptstyle (2)}(x) \right)
\end{align*}
\end{Lemma}

\begin{proof}
Let us start with the computation of $\text{Cov}(\mathcal{P}_{m}^{\scriptscriptstyle (1)}(t), \mathcal{P}_{m}^{\scriptscriptstyle (1)}(s))$, $i = 1$for $t \leq s$. 
\begin{align*}
\text{Cov}(\mathcal{P}_{m}^{\scriptscriptstyle (1)}(t), \mathcal{P}_{m}^{\scriptscriptstyle (1)}(s))& = \sum_{x, w_{1}(x) \in \mathbb{E}^{\scriptscriptstyle (1)}_{m}} \sum_{y, w_{1}(y) \in \mathbb{E}^{\scriptscriptstyle (1)}_{m}} \text{Cov}\left(f_{t}^{\scriptscriptstyle (1)}(x), f_{s}^{\scriptscriptstyle (1)}(y) \right).
\end{align*}
By stationarity, one has $\text{Cov}\left(f_{t}^{\scriptscriptstyle (1)}(x), f_{s}^{\scriptscriptstyle (1)}(y) \right)= \text{Cov}\left(f_{t}^{\scriptscriptstyle (1)}(0), f_{s}^{\scriptscriptstyle (1)}(y-x) \right)$, 
\begin{align*}
\text{Cov}\left(\mathcal{P}_{m}^{\scriptscriptstyle (1)}(t),\mathcal{P}_{m}^{\scriptscriptstyle (1)}(s) \right) & = \sum_{x_1 = 0}^{m-2}\sum_{x_2 = 0}^{m-1}\sum_{y_1 = 0}^{m-2}\sum_{y_2 = 0}^{m-1} \text{Cov}\left(f_{t}^{\scriptscriptstyle (1)}(0), f_{s}^{ \scriptscriptstyle (1)}(y-x) \right).
\end{align*}
For $x_1 \in [0,m-2]$ and $z_1 = y_1 - x_1$ one has
\begin{align*} 
\left\{
\begin{array}{rcl}
0 \leq x_1 \leq m-2 \\
0 \leq y_1 \leq m-2 
\end{array}\right.& \Leftrightarrow  \left\{ \begin{array}{rcl}
0 \leq x_1 \leq m-2 \hspace{0.7cm} \\
-x_1 \leq z_1 \leq m-2 - x_1
\end{array}\right. \\
& \Leftrightarrow  \left\{
\begin{array}{rcl}
2-m \leq z_1 \leq m-2 \hspace{2.5cm}  \\
\max(0, -z_1) \leq x_1 \leq \min(m-2 , m-2-z_1)
\end{array}\right..
\end{align*} 
Similarly for $z_2 = y_2 - x_2$ and $x_2 \in [0, m-1]$
\begin{align*}
 \left\{
\begin{array}{rcl}
0 \leq x_2 \leq m-1 \\
0 \leq y_2 \leq m-1 
\end{array}\right. & \Leftrightarrow
\left\{
\begin{array}{rcl}
1-m \leq z_2 \leq m-1 \hspace{2.5cm}  \\
\max(0, -z_2) \leq x_2 \leq \min(m-1 , m-1-z_2).
\end{array}\right., 
\end{align*}
Thus,
\begin{align*}
\text{Cov}\left(\mathcal{P}_{m}^{\scriptscriptstyle (1)}(t),\mathcal{P}_{m}^{\scriptscriptstyle (1)}(s) \right) &  = \sum_{z_{\scriptscriptstyle 1} = \left(2-m\right)}^{m-2}\sum_{z_{\scriptscriptstyle 2}=\left(1-m\right)}^{m-1}\left(m -1 -|z_{1}|\right)\left(m - |z_{2}|\right) \text{Cov}\left(f_{t}^{\scriptscriptstyle (1)}(0), f_{s}^{\scriptscriptstyle (1)}(z) \right) 
\end{align*}
For the second part of \eqref{covariance}, we have still by stationarity
\begin{align*}
\text{Cov}\left(\mathcal{P}_{m}^{\scriptscriptstyle (1)}(t),\mathcal{P}_{m}^{\scriptscriptstyle (2)}(s) \right) & = \sum_{x, w_{1}(x) \subset \text{\larger[1]{$($}}0, m \text{\larger[1]{$)$}}^{2}} \sum_{y, w_{2}(y) \subset \text{\larger[1]{$($}}0, m \text{\larger[1]{$)$}}^{2}} \text{Cov}\left(f_{t}^{\scriptscriptstyle (1)}(0), f_{s}^{ \scriptscriptstyle (2)}(y-x) \right)\\
& =  \sum_{x_1 = 0}^{m-2}\sum_{x_2 = 0}^{m-1}\sum_{y_1 = 0}^{m-1}\sum_{y_2 = 0}^{m-2}\text{Cov}\left(f_{t}^{\scriptscriptstyle (1)}(0), f_{s}^{ \scriptscriptstyle (2)}(y-x) \right)
\end{align*}
Again, considering $z_1 = y_1 - x_1$ and, $z_{2} = y_{2}-x_{2}$, we get
\begin{align*}
\text{Cov}\left(\mathcal{P}_{m}^{\scriptscriptstyle (1)}(t),\mathcal{P}_{m}^{\scriptscriptstyle (2)}(s) \right) & = \sum_{z_1=1}^{m-1}\sum_{z_2 = 0}^{m-2}\left(m-z_1\right)\left(m-z_2-1\right)\text{Cov}\left(f_{t}^{\scriptscriptstyle (1)}(0), f_{s}^{ \scriptscriptstyle (2)}(z) \right)\\
& +  \sum_{z_1=2-m}^{\scriptscriptstyle 0}\sum_{z_2 = 0}^{m-2}\left(m+z_1-1\right)\left(m-z_2-1\right)\text{Cov}\left(f_{t}^{\scriptscriptstyle (1)}(0), f_{s}^{ \scriptscriptstyle (2)}(z) \right)\\
& +  \sum_{z_1=1}^{m-1}\sum_{z_2 = -m+1}^{-1}\left(m-z_1\right)\left(m+z_2\right)\text{Cov}\left(f_{t}^{\scriptscriptstyle (1)}(0), f_{s}^{ \scriptscriptstyle (2)}(z) \right)\\
& + \sum_{z_1=2-m}^{\scriptscriptstyle 0}\sum_{z_2 = -m+1}^{-1} \left(m+z_1-1\right)\left(m+z_2\right)\text{Cov}\left(f_{t}^{\scriptscriptstyle (1)}(0), f_{s}^{ \scriptscriptstyle (2)}(z) \right) \\
& = \sum_{z_1=2-m}^{m-1} \sum_{z_2 = 1-m}^{m-2} \left(m-|z_1| - \mathbbm{1}_{z_{1} \leq 0}\right)\left(m-|z_2|-\mathbbm{1}_{z_{2} \geq 0}\right)\text{Cov}\left(f_{t}^{\scriptscriptstyle (1)}(0), f_{s}^{ \scriptscriptstyle (2)}(z) \right)
\end{align*}
\end{proof}
Getting the value of the covariance requires the computation of all the \linebreak elements~$\text{Cov}\left(f_{t}^{\scriptscriptstyle (j)}(0), f_{s}^{\scriptscriptstyle (i)}(x)\right)$, $i,j \in\{1,2\}$ which has a two, three and four cell structure. 
Figure \ref{configurations} below shows the five types of configurations depending on the orientation $i$ and position on the grid $x$. 
\begin{figure}[H]
\begin{center}
  \begin{subfigure}[b]{0.25\textwidth} 
  \centering
  \begin{tikzpicture}[scale = 1.5]
\draw[step = 0.5cm, black, thick](0,0) grid (1.5,1.5);
\draw[step = 2cm, line width =0.5mm, blue] (0.5,0) -- (0.5,0.5);
% \draw[step = 2cm, line width =0.5mm, blue] (1,0) -- (1,0.5);
\draw[pattern=north east lines, pattern color=blue] (0,0) rectangle (0.5,0.5);
\draw[pattern=north west lines, pattern color=green, opacity=1] (0, 0) rectangle (0.5,0.5);
\draw[pattern=north east lines, pattern color=red] (0.5, 0) rectangle (1,0.5);
\draw[pattern=north west lines, pattern color=cyan, opacity=1] (0.5, 0) rectangle (1,0.5);
\end{tikzpicture}
\caption{$i = 1, x = (0,0)$}
\end{subfigure}
\begin{subfigure}[b]{0.25\textwidth} 
  \centering
\begin{tikzpicture}[scale = 1.5]
\draw[step = 0.5cm, black, thick](0,0) grid (1.5,1.5);
\draw[step = 2cm, line width =0.5mm, blue] (0.5,0) -- (0.5,0.5);
\draw[step = 2cm, line width =0.5mm, blue] (1,0) -- (1,0.5);
\draw[pattern=north west lines, pattern color=blue] (0,0) rectangle (0.5,0.5);
\draw[pattern=north east lines, pattern color=red] (0.5, 0) rectangle (1,0.5);
\draw[pattern=north west lines, pattern color=cyan, opacity=1] (1, 0) rectangle (1.5,0.5);
\end{tikzpicture}
\caption{$i = 1$, $x = (0,1)$}
\end{subfigure}
\begin{subfigure}[b]{0.25\textwidth} 
  \centering 
\begin{tikzpicture}[scale = 1.5]
\draw[step = 0.5cm, black, thick](0,0) grid (1.5,1.5);
\draw[step = 2cm, line width =0.5mm, blue] (0.5,0) -- (0.5,0.5);
\draw[step = 2cm, line width =0.5mm, blue] (1,1) -- (1,1.5);
\draw[pattern=north west lines, pattern color=blue] (0,0) rectangle (0.5,0.5);
\draw[pattern=north west lines, pattern color=cyan, opacity=1] (0.5, 0) rectangle (1,0.5);
\draw[pattern=north west lines, pattern color=green, opacity=1] (1.5, 1.5) rectangle (1,1);
\draw[pattern=north west lines, pattern color=red, opacity=1] (0.5, 1)  rectangle (1, 1.5);
\end{tikzpicture}
\caption{$i = 1, x = (1,2)$}
\end{subfigure}
  \hspace{0.5cm}
\begin{subfigure}[b]{0.25\textwidth} 
\centering
\begin{tikzpicture}[scale =1.5]
\draw[step = 0.5cm, black, thick](0,0) grid (1.5,1.5);
\draw[step = 2cm, line width =0.5mm, blue] (0.5,0) -- (0.5,0.5);
\draw[step = 2cm, line width =0.5mm, blue] (0,0.5) -- (0.5,0.5);
\draw[pattern=north west lines, pattern color=blue] (0,0) rectangle (0.5,0.5);
\draw[pattern=north east lines, pattern color=red] (0, 0.5) rectangle (0.5,1);
\draw[pattern=north west lines, pattern color=cyan, opacity=1] (0.5, 0) rectangle (1,0.5);
\end{tikzpicture}
\caption{$i = 2, x = (0,0)$}
\end{subfigure}
\begin{subfigure}[b]{0.25\textwidth} 
\centering 
\begin{tikzpicture}[scale = 1.5]
\draw[step = 0.5cm, black, thick](0,0) grid (1.5,1.5);
\draw[step = 2cm, line width =0.5mm, blue] (0.5,0) -- (0.5,0.5);
\draw[step = 2cm, line width =0.5mm, blue] (1,1) -- (1.5,1);
\draw[pattern=north west lines, pattern color=blue] (0,0) rectangle (0.5,0.5);
\draw[pattern=north west lines, pattern color=cyan, opacity=1] (0.5, 0) rectangle (1,0.5);
\draw[pattern=north west lines, pattern color=green, opacity=1] (1.5, 1.5) rectangle (1,1);
\draw[pattern=north west lines, pattern color=red, opacity=1] (1, 0.5)  rectangle (1.5, 1);
\end{tikzpicture}
\caption{$i = 2, z = (2,1)$}
\end{subfigure}
\end{center}
\caption{Illustration of the two three and four cells configuration involved in the study of $\text{Cov}\left(f_{t}^{\scriptscriptstyle (1)}(0), f_{s}^{ \scriptscriptstyle (i)}(x)\right)$ in Lemma \ref{CovLemma}.}
\label{configurations}
\end{figure}
Let us first start with the two cell case, which appears in the computation of the term $\text{Cov}\left(f_{t}^{\scriptscriptstyle (1)}(0), f_{s}^{\scriptscriptstyle (1)}(0)\right)$, in order to compute it, we just have to apply the same strategy we used to compute the first moment in Proposition \ref{propFisrtmoment}. Thus, We get the next result. \\
\begin{Lemma}
\begin{align*}
\text{Cov}\left(f_{t}^{\scriptscriptstyle (1)}(0), f_{s}^{\scriptscriptstyle (1)}(0)\right) & = \mathbb{E}\left(\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle 0}, X_{\scriptscriptstyle  e_1}) \leq t \leq \max(X_{\scriptscriptstyle 0}, X_{e_1})\right\}}\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle 0}, X_{ \scriptscriptstyle  e_1}) \leq s \leq \max(X_{\scriptscriptstyle 0}, X_{\scriptscriptstyle e_1}) \right\}}\right) \\
& - \mathbb{E}\left(\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle 0}, X_{e_1}) \leq t \leq \max(X_{\scriptscriptstyle 0}, X_{e_1})\right\}}\right)\mathbb{E}\left(\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle 0}, X_{e_1}) \leq s \leq \max(X_{\scriptscriptstyle 0}, X_{e_1})\right\}}\right),
\end{align*}
where, 
{\small
\begin{align*}
\mathbb{E}\left(\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle 0}, X_{\scriptscriptstyle  e_1}) \leq t \leq \max(X_{\scriptscriptstyle 0}, X_{e_1})\right\}}\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle 0}, X_{ \scriptscriptstyle  e_1}) \leq s \leq \max(X_{\scriptscriptstyle 0}, X_{\scriptscriptstyle e_1}) \right\}}\right) & = \\
& \hspace{-9cm}\mathbb{E}\left(\left(\Phi\left(\dfrac{\sqrt{2}\max\left(t,s\right) + \sqrt{1-\rho(e_1)}|Z|}{\sqrt{1+\rho(e_1)}}\right)  - \Phi\left(\dfrac{\sqrt{2}\min\left(t,s\right) - \sqrt{1-\rho(e_1)}|Z|}{\sqrt{1+\rho(e_1)}}\right)\right)\right. \\
& \hspace{-1cm} \left.\mathbbm{1}_{\left\{(|s-t| \leq |Z|\sqrt{2(1-\rho(e_1))}\right\}}\right)
\end{align*}}
with $Z \sim \mathcal{N}(0,1)$. 
\end{Lemma}
Let us consider the Gaussian vector $\left(X_{\scriptscriptstyle 0}, X_{e_1}, X_{\scriptscriptstyle x}, X_{x+e_i}\right)$ of covariance matrix given by $$\Sigma_{i}(x) = \begin{pmatrix} 1 & \rho(e_1) & \rho(x) & \rho(x+e_i)\\
\rho(e_1) & 1 & \rho(x-e_1) &  \rho(x+e_i- e_1) \\ 
\rho(x) & \rho(x-e_1) & 1 & \rho(e_i) \\
\rho(x+e_i) & \rho(x+e_i-e_1) & \rho(e_i) & 1 \end{pmatrix}.$$ 
Let us define the following quantity  \begin{equation} \label{expressions} \Delta^{\scriptscriptstyle (i)}_{\scriptscriptstyle x} := X_{x+e_i} - X_{\scriptscriptstyle x}, \hspace{0.5cm} S^{\scriptscriptstyle (i)}_{\scriptscriptstyle x} := X_{x+e_i} + X_{\scriptscriptstyle x},\end{equation}
$\Delta \rho(x) = \rho(x-e_1)-\rho(x)$ and $S\rho(x)  =  \rho(x-e_1)+\rho(x)$. The covariance matrix of the Gaussian vector $\left(\Delta^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}, S^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}, \Delta^{\scriptscriptstyle (i)}_{\scriptscriptstyle x}, S^{\scriptscriptstyle (i)}_{\scriptscriptstyle x}\right)$ is equal to ${\tilde{\Sigma}_{i}(x)~= M\Sigma_{i}(x) M^{\star}}$ with {\small $M = \begin{pmatrix}
-1 & 1 & 0 &  0 \\
1 & 1 & 0 &  0 \\
0& 0 & -1 &  1 \\
0 & 0 & 1 & 1
\end{pmatrix}$}. Then $\tilde{\Sigma}_{i}(x) = \begin{pmatrix} A_{1} & B_{i}(x) \\ B^{\star}_{i}(x) & A_{i} \end{pmatrix}$
with 
$A_{i} = \begin{pmatrix} 2(1-\rho(e_i)) & 0 \\ 0 & 2(1+\rho(e_i))\end{pmatrix}$ and \linebreak {\small $B_{i}(x) =~\begin{pmatrix} \Delta \rho(x+e_i) - \Delta \rho(x)&    \Delta \rho(x) + \Delta \rho(x+e_i) \\ 
 S\rho(x+e_i)- S\rho(x) &  S\rho(x) + S\rho(x+e_i) 
 \end{pmatrix}$}. The quantity we aim to compute is 
{\small
\begin{align}
\label{covconf}
\text{Cov}\left(f_{t}^{\scriptscriptstyle (1)}(0), f_{s}^{ \scriptscriptstyle (i)}(x)\right) & = \nonumber \\
& \hspace{-2.7cm} \mathbb{E}\left(\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle 0}, X_{e_1}) \leq t \leq \max(X_{\scriptscriptstyle 0}, X_{e_1})\right\}}\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle x}, X_{x+e_i}) \leq s \leq \max(X_{\scriptscriptstyle x}, X_{x+e_i}) \right\}}\right) \nonumber\\
& \hspace{1cm} - \mathbb{E}\left(\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle 0}, X_{e_1}) \leq t \leq \max(X_{\scriptscriptstyle 0}, X_{e_1})\right\}}\right)\mathbb{E}\left(\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle 0}, X_{e_i}) \leq t \leq \max(X_{\scriptscriptstyle 0}, X_{e_i})\right\}}\right).
\end{align}}Considering the first part of the previous equation, and applying Lemma \ref{lemmasommediff}
{\small
\begin{align*}
\mathbb{E}\left(\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle 0}, X_{e_1}) \leq t \leq \max(X_{\scriptscriptstyle 0}, X_{e_1})\right\}}\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle x}, X_{x+e_i}) \leq s \leq \max(X_{\scriptscriptstyle x}, X_{x+e_i}) \right\}}\right)\\
& \hspace{-3cm}= \mathbb{E}\left(\mathbbm{1}_{\left\{ \left|2t - S^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}\right| \leq \left|\Delta^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}\right|\right\}}\mathbbm{1}_{\left\{ \left|2s - S^{\scriptscriptstyle (i)}_{\scriptscriptstyle x}\right| \leq \left|\Delta^{\scriptscriptstyle (i)}_x\right| \right\}}\right).
\end{align*}}Let us assume that $|\rho(e_1)| < 1$, the Cholesky decomposition of $\tilde{\Sigma}_{i}(x)$ allows us to get the expression of the Gaussian vector $\left(\Delta^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0},S^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0},\Delta^{\scriptscriptstyle (i)}_{\scriptscriptstyle x},S^{\scriptscriptstyle (i)}_{\scriptscriptstyle x} \right)$ in a new basis $\left(U,V,W,Z\right)$ with $\left(U, V, W, Z\right) \sim \mathcal{N}\left(0,I_{4}\right)$. We get the following decomposition \\ 
$\left\{
  \begin{array}{rlc}
  \Delta^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0} & = & \sigma^{\scriptscriptstyle (-)}_{1}U,  \textcolor{white}{+  \pi^{\scriptscriptstyle(+, i)}_{V}(x) V +  \pi^{\scriptscriptstyle(+, i)}_{W}(x) W +  \pi^{\scriptscriptstyle(+, i)}_{Z}(x)Z, } \\
  S^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0} & = & \textcolor{white}{\sigma^{\scriptscriptstyle (-)}_{1}U,} \sigma^{\scriptscriptstyle (+)}_{1}V, \textcolor{white}{+  \pi^{\scriptscriptstyle(+, i)}_{W}(x) W +  \pi^{\scriptscriptstyle(+, i)}_{Z}(x)Z, } \\
  \Delta^{\scriptscriptstyle (i)}_{\scriptscriptstyle  x}& = &  \pi^{\scriptscriptstyle(-, i)}_{U}(x)U +  \pi^{\scriptscriptstyle(-, i)}_{V}(x) V +  \pi^{\scriptscriptstyle(-, i)}_{W}(x) W, \textcolor{white}{ + \pi^{\scriptscriptstyle(+, i)}_{Z}(x)Z, } \\
  S^{\scriptscriptstyle (i)}_{\scriptscriptstyle x} & = &    \pi^{\scriptscriptstyle(+, i)}_{U}(x)U +  \pi^{\scriptscriptstyle(+, i)}_{V}(x) V +  \pi^{\scriptscriptstyle(+, i)}_{W}(x) W +  \pi^{\scriptscriptstyle(+, i)}_{Z}(x)Z, 
\end{array}\right.$ \\
with 
$\left\{
 \begin{array}{rlc}
 \sigma^{\scriptscriptstyle (-)}_{1} & = &  \sqrt{2(1-\rho(e_1))} \\
 \sigma^{\scriptscriptstyle (+)}_{1} & = & \sqrt{2(1+\rho(e_1))} \\
 \end{array}\right.$ 
 $\left\{
 \begin{array}{rlc}
   \pi^{\scriptscriptstyle(-, i)}_{U}(x) & = & \frac{1}{\sigma^{\scriptscriptstyle (-)}_{1}}\left(\Delta \rho(x+e_i) -\Delta \rho(x)\right) \\
   \pi^{\scriptscriptstyle(-, i)}_{V}(x) & = &  \frac{1}{\sigma^{\scriptscriptstyle (+)}_{1}}\left(S\rho(x+e_i)-S\rho(x)\right)  \\
   \pi^{\scriptscriptstyle(-, i)}_{W}(x)& = & \left(2(1 - \rho(e_{i})) - \pi^{\scriptscriptstyle(-, i)}_{U}(x)^{2} - \pi^{\scriptscriptstyle(-, i)}_{V}(x)^{2}\right)^{\scriptscriptstyle (1/2)} \end{array}\right.$ \\
    $\left\{
 \begin{array}{rlc}
   \pi^{\scriptscriptstyle(+, i)}_{U}(x) & = &  \frac{1}{\sigma^{\scriptscriptstyle (-)}_{1}}\left(\Delta \rho(x) + \Delta \rho(x+e_i)\right) \\
    \pi^{\scriptscriptstyle(+, i)}_{V}(x) & = & \frac{1}{\sigma^{\scriptscriptstyle (+)}_{1}}\left(S \rho(x) + S\rho(x+e_i)\right) \\
     \pi^{\scriptscriptstyle(+, i)}_{W}(x) & = & \frac{-1}{\pi^{\scriptscriptstyle(-, i)}_{W}(x)}\left(\pi^{\scriptscriptstyle(-, i)}_{U}(x)\pi^{\scriptscriptstyle(+, i)}_{U}(x) + \pi^{\scriptscriptstyle(-, i)}_{V}(x)\pi^{\scriptscriptstyle(+, i)}_{V}(x)\right) \\
   \pi^{\scriptscriptstyle(+, i)}_{Z}(x)& = & \left(2(1 + \rho(e_i)) - \pi^{\scriptscriptstyle(+, i)}_{U}(x)^{2} - \pi^{\scriptscriptstyle(+, i)}_{V}(x)^{2} - \pi^{\scriptscriptstyle(+, i)}_{W}(x)^{2}\right)^{\scriptscriptstyle (1/2)}
\end{array}\right.$ 
\begin{Proposition}
\label{PropositionFourCell}
Considering the assumptions above we get 
{\small 
\begin{align}
\label{Thisthofour}
\mathbb{E}\left(\mathbbm{1}_{\left\{ \left|2t - S^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}\right| \leq \left|\Delta^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}\right|\right\}}\mathbbm{1}_{\left\{ \left|2s - S^{\scriptscriptstyle i}_{\scriptscriptstyle x}\right| \leq \left|\Delta^{\scriptscriptstyle (i)}_x\right| \right\}}\right) & = \mathbb{E}\left(\mathbbm{1}_{\left\{ |t - \frac{\sigma^{\scriptscriptstyle (+)}_{1}V}{2}| \leq |\frac{\sigma^{\scriptscriptstyle (-)}_{1}U}{2}|\right\}}\right. \nonumber\\
& \hspace{-6cm} \left. \left|\Phi_{ \pi^{\scriptscriptstyle(+, i)}_{Z}(x)}\left( 2s - \frac{2\Delta \rho(x) }{\sigma^{\scriptscriptstyle (-)}_{1}} U - \frac{2S\rho(x)}{\sigma^{\scriptscriptstyle (+)}_{1}}V - \left( \pi^{\scriptscriptstyle(+, i)}_{W}(x) -  \pi^{\scriptscriptstyle(-, i)}_{W}(x)\right) W \right) \right. \right.  \nonumber \\
& \hspace{-4cm}\left. \left.  - \Phi_{ \pi^{\scriptscriptstyle(+, i)}_{Z}(x)}\left( 2s - \frac{2\Delta \rho(x+e_i) }{\sigma^{\scriptscriptstyle (-)}_{1}}U - \frac{2S \rho(x+e_i) }{\sigma^{\scriptscriptstyle (+)}_{1}}V - \left( \pi^{\scriptscriptstyle(+, i)}_{W}(x) +  \pi^{\scriptscriptstyle(-, i)}_{W}(x) \right)W \right)  \right| \right)
\end{align}
}with $\Phi_{\pi^{\scriptscriptstyle(+, i)}_{Z}(x)}$ the cumulative distribution function of $Z \sim \mathcal{N}(0, (\pi^{\scriptscriptstyle(+, i)}_{Z}(x))^{2})$.
\end{Proposition}
\begin{proof}(Proof of Proposition \ref{PropositionFourCell})
Let us first consider the following indicator function
{\small
\begin{align*}
\mathbbm{1}_{\left\{ |2s - S^{i}_{\scriptscriptstyle x}| \leq |\Delta^{\scriptscriptstyle(i)}_ {\scriptscriptstyle x}| \right\}}
& = \mathbbm{1}_{\big\{ 2s -  \pi^{\scriptscriptstyle(+, i)}_{U}(x)U -  \pi^{\scriptscriptstyle(+, i)}_{V}(x) V -  \pi^{\scriptscriptstyle(+, i)}_{W}(x) W  - |\Delta^{\scriptscriptstyle(i)}_ {\scriptscriptstyle x}| \leq \pi^{\scriptscriptstyle(+, i)}_{Z}(x)Z} \\ 
& \hspace{3cm} _{\leq 2s -  \pi^{\scriptscriptstyle(+, i)}_{U}(x)U -  \pi^{\scriptscriptstyle(+, i)}_{V}(x) V -  \pi^{\scriptscriptstyle(+, i)}_{W}(x) W + |\Delta^{\scriptscriptstyle(i)}_ {\scriptscriptstyle x}| \big\}}
\end{align*}}Thus, 
\begin{align*}
\mathbb{E}\left(\mathbbm{1}_{\left\{ |2t - S^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}| \leq |\Delta^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}|\right\}}\mathbbm{1}_{\left\{ |2s - S^{\scriptscriptstyle (i)}_{\scriptscriptstyle x}| \leq |\Delta^{\scriptscriptstyle (i)}_ {\scriptscriptstyle x}| \right\}}\right)\\
&\hspace{-6.5cm} = \mathbb{E}\left(\mathbbm{1}_{\left\{ |2t - S^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}| \leq |\Delta^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}|\right\}}\mathbbm{1}_{\big\{ 2s -  \pi^{\scriptscriptstyle(+, i)}_{U}(x)U -  \pi^{\scriptscriptstyle(+, i)}_{V}(x) V -  \pi^{\scriptscriptstyle(+, i)}_{W}(x) W  - |\Delta^{\scriptscriptstyle(i)}_ {\scriptscriptstyle x}| \leq \pi^{\scriptscriptstyle(+, i)}_{Z}(x)Z} \right.  \\  
& \hspace{3cm}\left. _{\leq 2s -  \pi^{\scriptscriptstyle(+, i)}_{U}(x)U -  \pi^{\scriptscriptstyle(+, i)}_{V}(x) V -  \pi^{\scriptscriptstyle(+, i)}_{W}(x) W + |\Delta^{\scriptscriptstyle(i)}_ {\scriptscriptstyle x}| \big\}}\right)\\
& \hspace{-6.5cm} \text{The variable}\; Z \sim \mathcal{N}\left(0, \pi^{\scriptscriptstyle(+, i)}_{Z}(x)^{2}\right) \text{and} \; \text{$\Phi_{\pi^{\scriptscriptstyle(+, i)}_{Z}(x)}$ being an increasing function,}\\
& \hspace{-6.5cm} = \mathbb{E}\left(\mathbbm{1}_{\left\{ |t - \frac{\sigma^{\scriptscriptstyle (+)}_{1}V}{2}| \leq |\frac{\sigma^{\scriptscriptstyle (-)}_{1}U}{2}|\right\}}\right. \\
&\hspace{-4cm} \left. \left|\Phi_{ \pi^{\scriptscriptstyle(+, i)}_{Z}(x)}\left( 2s -  \pi^{\scriptscriptstyle(+, i)}_{U}(x)U -  \pi^{\scriptscriptstyle(+, i)}_{V}(x) V -  \pi^{\scriptscriptstyle(+, i)}_{W}(x) W  + \Delta^{\scriptscriptstyle(i)}_ {\scriptscriptstyle x}  \right) \right. \right.  \nonumber \\
& \hspace{-4cm}\left. \left.  - \Phi_{ \pi^{\scriptscriptstyle(+, i)}_{Z}(x)}\left( 2s -  \pi^{\scriptscriptstyle(+, i)}_{U}(x)U -  \pi^{\scriptscriptstyle(+, i)}_{V}(x) V -  \pi^{\scriptscriptstyle(+, i)}_{W}(x) W  - \Delta^{\scriptscriptstyle(i)}_ {\scriptscriptstyle x} \right)  \right| \right) \\
& \hspace{-6.5cm} = \mathbb{E}\left(\mathbbm{1}_{\left\{ |t - \frac{\sigma^{\scriptscriptstyle (+)}_{1}V}{2}| \leq |\frac{\sigma^{\scriptscriptstyle (-)}_{1}U}{2}|\right\}}\right. \\
&\hspace{-6.2cm} \left. \left|\Phi_{ \pi^{\scriptscriptstyle(+, i)}_{Z}(x)}\left( 2s + \left( \pi^{\scriptscriptstyle(-, i)}_{U}(x)- \pi^{\scriptscriptstyle(+, i)}_{U}(x)\right) U + \left(\pi^{\scriptscriptstyle(-, i)}_{V}(x)- \pi^{\scriptscriptstyle(+, i)}_{V}(x)\right)V + \left(\pi^{\scriptscriptstyle(-, i)}_{W}(x) - \pi^{\scriptscriptstyle(+, i)}_{W}(x)\right) W \right) \right. \right.  \nonumber \\
& \hspace{-6.5cm}\left. \left.  - \Phi_{ \pi^{\scriptscriptstyle(+, i)}_{Z}(x)}\left( 2s - \left( \pi^{\scriptscriptstyle(-, i)}_{U}(x)+ \pi^{\scriptscriptstyle(+, i)}_{U}(x)\right) U - \left(\pi^{\scriptscriptstyle(-, i)}_{V}(x)+ \pi^{\scriptscriptstyle(+, i)}_{V}(x)\right)V   - \left( \pi^{\scriptscriptstyle(+, i)}_{W}(x) +  \pi^{\scriptscriptstyle(-, i)}_{W}(x) \right)W \right)  \right| \right)
\end{align*}
\end{proof}
\begin{remark} In the third cell configuration,the matrix $\Sigma_{i}(x)$ is not positive-definite, thus the Cholesky decomposition does not apply, we have to proceed a differently, but this case is not relevant to the establishment of the Central Limit Theorem, therefor its computations will be presented in the Appendix.  
\end{remark}
\subsubsection{Spatial independent case} 
Let us start by considering the four cell configuration and let us see what happens when ${\rho(x) = \rho(x+e_i) = \rho(x-e_i) = \rho(x + e_i - e_1) = 0}$, supposedly when $x > > 0$. Then
$\pi^{\scriptscriptstyle(-, i)}_{W}(x) = \sqrt{2(1-\rho(e_i))}$, $\pi^{\scriptscriptstyle(+, i)}_{W}(x) = 0$ and $\pi^{\scriptscriptstyle(+, i)}_{Z}(x) = \sqrt{2(1+\rho(e_i))}.$ Thus, Equation \eqref{Thistho} becomes 
{\small
\begin{align*}
\mathbb{E}\left(\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle 0}, X_{e_1}) \leq t \leq \max(X_{\scriptscriptstyle 0}, X_{e_1})\right\}}\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle x}, X_{x+e_i}) \leq s \leq \max(X_{\scriptscriptstyle x}, X_{x+e_i}) \right\}}\right) \nonumber \\
& \hspace{-10cm} = \mathbb{E}\left(\mathbbm{1}_{\left\{ |t - \frac{\sqrt{2(1+\rho(e_1))}V}{2}| \leq |\frac{\sqrt{2(1-\rho(e_1))}U}{2}|\right\}} \right.  \\
& \left. \hspace{-10cm} \left(\mathbbm{1}_{\left\{\sqrt{2(1-\rho(e_i))}W \geq 0\right\}} \left(\Phi_{\pi^{\scriptscriptstyle(+, i)}_{Z}(x)}\left( 2s + \sqrt{2(1-\rho(e_1))} W\right) - \Phi_{\pi^{\scriptscriptstyle(+, i)}_{Z}(x)}\left(2s - \sqrt{2(1-\rho(e_1))} W\right)\right) \right. \right. \\
&  \hspace{-10cm} \left. \left. \mathbbm{1}_{\left\{\sqrt{2(1-\rho(e_1))} W  \leq 0\right\}} \left(\Phi_{\pi^{\scriptscriptstyle(+, i)}_{Z}(x)}\left( 2s - \sqrt{2(1-\rho(e_1))}W \right) - \Phi_{\pi^{\scriptscriptstyle(+, i)}_{Z}(x)}\left(2s + \sqrt{2(1-\rho(e_1))} W\right)\right)\right)\right)\\ 
& \hspace{-10cm} = \mathbb{E}\left(\mathbbm{1}_{\left\{ |t - \frac{\sqrt{2(1+\rho(e_1))}V}{2}| \leq |\frac{\sqrt{2(1-\rho(e_1))}U}{2}|\right\}}\mathbbm{1}_{\left\{ |s - \frac{\sqrt{2(1+\rho(e_i))}Z}{2}| \leq |\frac{\sqrt{2(1-\rho(e_i))}W}{2}|\right\}}\right) \\
& \hspace{-10cm} = \mathbb{E}\left(\mathbbm{1}_{\left\{ |t - \frac{\sqrt{2(1+\rho(e_1))}V}{2}| \leq |\frac{\sqrt{2(1-\rho(e_1))}U}{2}|\right\}}\right)\mathbb{E}\left(\mathbbm{1}_{\left\{ |s - \frac{\sqrt{2(1+\rho(e_i))}Z}{2}| \leq |\frac{\sqrt{2(1-\rho(e_i))}W}{2}|\right\}}\right).
\end{align*}
}Thus, $\text{Cov}\left(f_{t}^{\scriptscriptstyle (1)}(0), f_{s}^{ \scriptscriptstyle (j)}(z)\right) = 0.$
\subsubsection{Behavior at large distance.}
Considering the case where the size of the window $S$ goes to infinity, and assuming that $\rho(x) \to 0$ when $||x||_{2} \to \infty$, we would like to study the decay of $\text{Cov}\left(f_{t}^{\scriptscriptstyle(1)}(0), f_{s}^{\scriptscriptstyle(i)}(x) \right)$ for~$||x||_{2} >> 1$,  
\begin{Lemma}
Let $r(x) = \max\left(|\rho(x)|, |\rho(x + e_i)|, |\rho(x-e_1)|, |\rho(x +e_i - e_1)| \right)$, there exist a constant $C \in \mathbb{R}^{+}$, for all $x \in \mathbb{G}_{m}$
\begin{align*}
&  |\text{Cov}\left(f_{t}^{\scriptscriptstyle(1)}(0), f_{s}^{\scriptscriptstyle(i)}(x) \right)| \leq Cr(x). 
\end{align*}
Besides, if $\sum_{x \in \mathbb{Z}^{2}} r(x) <~\infty$, then 
$\lim_{m \to \infty}\text{Cov}\left(\frac{1}{m}\mathcal{P}_{m}^{\scriptscriptstyle (1)}(t), \frac{1}{m}\mathcal{P}_{m}^{\scriptscriptstyle (i)}(s) \right)$ exists. 
\end{Lemma}
\begin{proof}
Let us begin by reconsidering the previous quantities and 
\begin{table}[H]
\centering 
\begin{tabularx}{\textwidth}{|P{3.5cm}|P{11.85cm}|}
\hline
Variable & upper bound \\ \hline
$\left|\pi^{\scriptscriptstyle(-, i)}_{U}(x)\right|$ & $\frac{4r(x)}{\sigma^{\scriptscriptstyle (-)}_{1}}$ \\  \hline
$\left|\pi^{\scriptscriptstyle(-, i)}_{V}(x)\right|$ & $\frac{4r(x)}{\sigma^{\scriptscriptstyle (+)}_{1}}$  \\  \hline
$\left|\pi^{\scriptscriptstyle(-, i)}_{W}(x) - \sigma^{(-)}_{i}\right|$ & $ \frac{16r^{2}(x)}{\sigma^{(-)}_{i}}\left(\frac{1}{(\sigma^{(-)}_{i})^{2}} + \frac{1}{(\sigma^{(+)}_{i})^{2}} \right)$ \\ 
\hline
$|\pi^{\scriptscriptstyle(+, i)}_{U}(x)|$ &  $\frac{4r(x)}{\sigma^{\scriptscriptstyle (-)}_{1}}$   \\ \hline
$|\pi^{\scriptscriptstyle(+, i)}_{V}(x)|$ &   $\frac{4r(x)}{\sigma^{\scriptscriptstyle (+)}_{1}}$  \\ \hline 
$ \left|\pi^{\scriptscriptstyle(+, i)}_{W}(x)\right| $ & $16r(x)^{2}\left(\sigma^{(-)}_{i} -  \frac{16r^{2}(x)}{\sigma^{(-)}_{i}}\left(\frac{1}{(\sigma^{(-)}_{i})^{2}} + \frac{1}{(\sigma^{(+)}_{i})^{2}} \right)\right)^{-1}\left(\frac{1}{(\sigma^{(-)}_{i})^{2}} + \frac{1}{(\sigma^{(+)}_{i})^{2}} \right)$\\ \hline
$\left|\pi^{\scriptscriptstyle(+, i)}_{Z}(x) - \sigma^{(+)}_{i}\right|$ & $\frac{16r^{2}(x)}{\sigma^{(+)}_{i}}\left(\frac{1}{(\sigma^{(-)}_{i})^{2}} + \frac{1}{(\sigma^{(+)}_{i})^{2}} +  \frac{\pi^{\scriptscriptstyle(+, i)}_{W}(x)^{2}}{16r^{2}(x)} \right)$\\\hline
\end{tabularx}
\end{table}
Let, $R_{1}(x) = \frac{2s + \left( \pi^{\scriptscriptstyle(-, i)}_{U}(x)- \pi^{\scriptscriptstyle(+, i)}_{U}(x)\right) U + \left(\pi^{\scriptscriptstyle(-, i)}_{V}(x)- \pi^{\scriptscriptstyle(+, i)}_{V}(x)\right)V + \left(\pi^{\scriptscriptstyle(-, i)}_{W}(x) - \pi^{\scriptscriptstyle(+, i)}_{W}(x)\right) W}{\pi^{\scriptscriptstyle(+, i)}_{Z}(x)} - \frac{2s + \sigma^{\scriptscriptstyle (-)}_{i}W}{\sigma^{\scriptscriptstyle (+)}_{i}}$ and $ R_{2}(x) = \frac{2s - \left( \pi^{\scriptscriptstyle(-, i)}_{U}(x)+\pi^{\scriptscriptstyle(+, i)}_{U}(x)\right) U - \left(\pi^{\scriptscriptstyle(-, i)}_{V}(x)+\pi^{\scriptscriptstyle(+, i)}_{V}(x)\right)V - \left(\pi^{\scriptscriptstyle(-, i)}_{W}(x) +\pi^{\scriptscriptstyle(+, i)}_{W}(x)\right) W}{\pi^{\scriptscriptstyle(+, i)}_{Z}(x)} - \frac{2s - \sigma^{\scriptscriptstyle (-)}_{i}W}{\sigma^{\scriptscriptstyle (+)}_{i}} $. Considering $R_{1}(x)$
{\small
\begin{align*}
|R_{1}(x)| = \left|\frac{\sigma^{(+)}_{i}\left(2s + \pi^{\scriptscriptstyle(-, i)}_{W}(x) W\right) - \pi^{\scriptscriptstyle(+, i)}_{Z}(x)\left( 2s +  \sigma^{(-)}_{i}W\right)}{\sigma^{(+)}_{i}\pi^{\scriptscriptstyle(+, i)}_{Z}(x)} \right.\\
& \hspace{-5cm} \left. + \frac{\sigma^{(+)}_{i}\left(\left( \pi^{\scriptscriptstyle(-, i)}_{U}(x)- \pi^{\scriptscriptstyle(+, i)}_{U}(x)\right) U + \left(\pi^{\scriptscriptstyle(-, i)}_{V}(x)- \pi^{\scriptscriptstyle(+, i)}_{V}(x)\right)V - \pi^{\scriptscriptstyle(+, i)}_{W}(x) W \right)}{\sigma^{(+)}_{i}\pi^{\scriptscriptstyle(+, i)}_{Z}(x)}\right| \\
& \hspace{-10cm} \text{making use of the previous inequalities, there exist a constant $C_{1} \in \mathbb{R}^{+}$ depending}\\
& \hspace{-10cm}  \text{on $s, \sigma^{(+)}_{i}, \sigma^{(-)}_{i}$ such that,} \\
& \hspace{-8.8cm}\leq C_{1}r(x)\left(|U| + |V| + |W|\right)
\end{align*}}
Also, there exist $C_2$ such that $R_{2}(x) \leq C_{2}r(x)\left(|U| + |V| + |W|\right)$. 
Reconsidering Equation~(\ref{Thisthofour}) we get,
{\small
\begin{align*}
 \text{Cov}\left(f_{t}^{\scriptscriptstyle(1)}(0), f_{s}^{\scriptscriptstyle(i)}(x) \right) & =  \\ 
 & \hspace{-2.5cm} \mathbb{E}\left(\mathbbm{1}_{\left\{ |t - \frac{\sigma^{\scriptscriptstyle (+)}_{1}V}{2}| \leq |\frac{\sigma^{\scriptscriptstyle (-)}_{1}U}{2}|\right\}}\right. \\
&\hspace{-2.5cm} \left. \left|\Phi_{ \pi^{\scriptscriptstyle(+, i)}_{Z}(x)}\left( 2s + \left( \pi^{\scriptscriptstyle(-, i)}_{U}(x)- \pi^{\scriptscriptstyle(+, i)}_{U}(x)\right) U + \left(\pi^{\scriptscriptstyle(-, i)}_{V}(x)- \pi^{\scriptscriptstyle(+, i)}_{V}(x)\right)V + \left(\pi^{\scriptscriptstyle(-, i)}_{W}(x) - \pi^{\scriptscriptstyle(+, i)}_{W}(x)\right) W \right) \right. \right.  \nonumber \\
& \hspace{-2.5cm}\left. \left.  - \Phi_{ \pi^{\scriptscriptstyle(+, i)}_{Z}(x)}\left( 2s - \left( \pi^{\scriptscriptstyle(-, i)}_{U}(x)+ \pi^{\scriptscriptstyle(+, i)}_{U}(x)\right) U - \left(\pi^{\scriptscriptstyle(-, i)}_{V}(x)+ \pi^{\scriptscriptstyle(+, i)}_{V}(x)\right)V   - \left( \pi^{\scriptscriptstyle(+, i)}_{W}(x) +  \pi^{\scriptscriptstyle(-, i)}_{W}(x) \right)W \right)  \right| \right) \\
& \hspace{-2.5cm} - \mathbb{E}\left(\mathbbm{1}_{\left\{ |t - \frac{\sigma^{\scriptscriptstyle (+)}_{1}V}{2}| \leq |\frac{\sigma^{\scriptscriptstyle (-)}_{1}U}{2}|\right\}}\left|\Phi\left(\frac{2s + \sigma^{\scriptscriptstyle (-)}_{i}W}{\sigma^{\scriptscriptstyle (+)}_{i}} \right) - \Phi\left(\frac{2s -\sigma^{\scriptscriptstyle (-)}_{i}W }{\sigma^{\scriptscriptstyle (+)}_{i}} \right)\right|\right)\\
& \hspace{-2.5cm} = \mathbb{E}\left(\mathbbm{1}_{\left\{ |t - \frac{\sigma^{\scriptscriptstyle (+)}_{1}V}{2}| \leq |\frac{\sigma^{\scriptscriptstyle (-)}_{1}U}{2}|\right\}} \left(\left|\Phi\left(\frac{2s + \sigma^{\scriptscriptstyle (-)}_{i}W}{\sigma^{\scriptscriptstyle (+)}_{i}} + R_{1}(x)) \right) - \Phi\left(\frac{2s -\sigma^{\scriptscriptstyle (-)}_{i}W }{\sigma^{\scriptscriptstyle (+)}_{i}} + R_{2}(x))\right)\right| \right. \right. \\
& \hspace{4.75cm} \left.\left. -  \left|\Phi\left(\frac{2s + \sigma^{\scriptscriptstyle (-)}_{i}W}{\sigma^{\scriptscriptstyle (+)}_{i}} \right) - \Phi\left(\frac{2s -\sigma^{\scriptscriptstyle (-)}_{i}W }{\sigma^{\scriptscriptstyle (+)}_{i}} \right)\right| \right)\right),
\end{align*}
}
thus, 
{\small
\begin{align*}
 |\text{Cov}\left(f_{t}^{\scriptscriptstyle(1)}(0), f_{s}^{\scriptscriptstyle(i)}(x) \right)| & \leq \\
 &  \mathbb{E}\left(\left|\Phi\left(\frac{2s + \sigma^{\scriptscriptstyle (-)}_{i}W}{\sigma^{\scriptscriptstyle (+)}_{i}} + R_{1}(x) \right)  - \Phi\left(\frac{2s + \sigma^{\scriptscriptstyle (-)}_{i}W}{\sigma^{\scriptscriptstyle (+)}_{i}} \right) \right. \right. \\
 & \hspace{2cm} \left. \left.+ \Phi\left(\frac{2s - \sigma^{\scriptscriptstyle (-)}_{i}W}{\sigma^{\scriptscriptstyle (+)}_{i}} \right) - \Phi\left(\frac{2s -\sigma^{\scriptscriptstyle (-)}_{i}W }{\sigma^{\scriptscriptstyle (+)}_{i}} + R_{2}(x))\right)\right|\right) \\
 & \leq \mathbb{E}\left(\left|\Phi\left(\frac{2s + \sigma^{\scriptscriptstyle (-)}_{i}W}{\sigma^{\scriptscriptstyle (+)}_{i}} + |R_{1}(x)| \right)  - \Phi\left(\frac{2s + \sigma^{\scriptscriptstyle (-)}_{i}W}{\sigma^{\scriptscriptstyle (+)}_{i}} \right) \right| \right. \\
 & \hspace{2cm} \left. + \left| \Phi\left(\frac{2s - \sigma^{\scriptscriptstyle (-)}_{i}W}{\sigma^{\scriptscriptstyle (+)}_{i}} \right) - \Phi\left(\frac{2s -\sigma^{\scriptscriptstyle (-)}_{i}W }{\sigma^{\scriptscriptstyle (+)}_{i}} + R_{2}(x))\right)\right|\right),
 \end{align*}}
Applying the mean value theorem, there exist a constant $C \in \mathbb{R}^{+}$ such that 
\begin{align*}
 |\text{Cov}\left(f_{t}^{\scriptscriptstyle(1)}(0), f_{s}^{\scriptscriptstyle(i)}(x) \right)| & \leq C\mathbb{E}\left(|R_{1}(x)| + |R_{2}(x)|\right) \\
& \leq Cr(x)
\end{align*}
Considering the series $\sum_{x \in \mathbb{Z}^{2}}\left(\mathbbm{1}_{x_{1} \in [2-m, m-2]}\mathbbm{1}_{x_{2} \in [1-m, m-1]} \frac{(m-|x_2|)(m-|x_1| -1)}{m^2} \text{Cov}\left(f_{t}^{\scriptscriptstyle(1)}(0), f_{s}^{\scriptscriptstyle(i)}(x) \right) \right)$
The limit of the general term as $m$ goes to infinity is
\begin{align*}
\lim_{m \to \infty}\frac{(m-|x_2|)(m-|x_1| -1)}{m^2} \text{Cov}\left(f_{t}^{\scriptscriptstyle(1)}(0), f_{s}^{\scriptscriptstyle(i)}(x) \right)\mathbbm{1}_{x_{1} \in [2-m, m-2]\cap \mathbb{Z}}\mathbbm{1}_{x_{2} \in [1-m, m-1]\cap \mathbb{Z}} \\
& \hspace{-7cm} =\text{Cov}\left(f_{t}^{\scriptscriptstyle(1)}(0), f_{s}^{\scriptscriptstyle(i)}(x) \right)\mathbbm{1}_{x \in \mathbb{Z}^{2}} 
\end{align*}
Besides, we just proved that for $m \in \mathbb{N}$, for all $x\in \mathbb{Z}^{2}$, $\text{Cov}\left(f_{t}^{\scriptscriptstyle(1)}(0), f_{s}^{\scriptscriptstyle(i)}(x)\right) \leq Cr(x)$, thus applying the dominated convergence theorem, $\lim_{m \to \infty}\text{Cov}\left(\frac{1}{m}\mathcal{P}_{m}^{\scriptscriptstyle (1)}(t), \frac{1}{m}\mathcal{P}_{m}^{\scriptscriptstyle (i)}(s) \right)$ exists.
\end{proof}
Now that we know more about how the covariance behaves, we can construct a Multi-directional multivariate CLT that will lead us to an isotropy test.
\vspace{-0.5cm}{}
\subsection{Isotropy test using the oriented perimeter}
In order to get the CLT we have to add some assumptions on the covariance structure of the field, we will need the following  assumptions 
\begin{itemize}
\item \textbf{$(A_{1})$} $\lim_{n \to \infty} m^{-2}\sum_{x, y \in \mathbb{G}_{m}}\rho(x-y)$ exists.
\item \textbf{$(A_{2})$} $\lim_{n \to \infty} m^{-2}\sum_{x, y \in \mathbb{G}_{m}}\rho(x-~y)^{2}$ exists.
\end{itemize} 
\begin{Theorem}\label{TCL}(Multi-directional multivariate CLT for r-threshold). We consider taht $X_{x \in \mathbb{G}_{m}}$ a centered stationary standard Gaussian, let $r$ be a positive integer, $m \geq 2$ and $t_{1}, \cdots,t_{r} \in \mathbb{R}$, then, under \textbf{$(A_{1})$} and \textbf{$(A_{2})$},
\begin{equation*}
\frac{1}{m}\left(\begin{pmatrix} \mathcal{P}^{\scriptscriptstyle (1)}_{m}(t_{1}) \\  \vdots \\ \mathcal{P}^{\scriptscriptstyle (1)}_{m}(t_{r}) \\ \mathcal{P}^{\scriptscriptstyle (2)}_{m}(t_{2}) \\  \vdots \\ \mathcal{P}^{\scriptscriptstyle (2)}_{m}(t_{r})  \end{pmatrix} - \begin{pmatrix} \mathbb{E}(\mathcal{P}^{\scriptscriptstyle (1)}_{m}(t_{1}))  \\ \vdots \\ \mathbb{E}(\mathcal{P}^{\scriptscriptstyle (1)}_{m}(t_{r}))\\\mathbb{E}(\mathcal{P}^{\scriptscriptstyle (2)}_{m}(t_{1})) \\ \vdots \\ \mathbb{E}(\mathcal{P}^{\scriptscriptstyle (2)}_{m}(t_{r})) \end{pmatrix}\right) \xrightarrow[m \to \infty]{d} \mathcal{N}\left(0,\,\Sigma_{2r}^{\star}\right),
\end{equation*}
where $\xrightarrow[]{d}$ holds for the convergence in distribution and $\mathcal{N}\left(0,\,\Sigma_{2r}^{\star}\right)$ for the $2r$-dimensional centered Gaussian distribution with covariance matrix $\Sigma^\star_{2r}$ given by
\begin{align*}
\Sigma^{\star}_{2r}(l,i,k,j) & : = \lim_{m \to \infty}\text{Cov}\left(\frac{1}{m}\mathcal{P}^{\scriptscriptstyle (i)}_{m}(t_{l}),  \frac{1}{m}\mathcal{P}^{\scriptscriptstyle (j)}_{m}(t_{k}) \right)
\end{align*}
\end{Theorem}
\begin{proof} For the sake of simplicity, we will prove the CLT for $r=1$, the generalised case is done following the same procedure. Following the idea that has been presented in \cite{Psymetrie}, 
using the Cramèr-Wold method, we need to prove that for each $(a_1, a_2) \in \mathbb{R}^{2}$, we have 
\begin{align*}
\frac{a_{1}}{m}\left(\mathcal{P}^{\scriptscriptstyle (1)}(t) - \mathbb{E}\left(\mathcal{P}^{\scriptscriptstyle (1)}(t)\right)\right) + \frac{a_{2}}{m}\left(\mathcal{P}^{\scriptscriptstyle (2)}(t) - \mathbb{E}\left(\mathcal{P}^{\scriptscriptstyle (2)}(t)\right)\right) \xrightarrow[m \to \infty]{d} \mathcal{N}\left(0,\,\sigma^{2}\right) 
\end{align*}
with $\sigma^{2} = \lim_{m \to \infty} \mathbb{V}\left( \frac{a_{1}}{m}\mathcal{P}^{\scriptscriptstyle (1)}(t) + \frac{a_{2}}{m}\mathcal{P}^{\scriptscriptstyle (2)}(t) \right)$.
We first need to recall some notations that were introduced in the first section \eqref{fi}. Let $\mathbb{G}_{m} := \text{\larger[1]{$[$}}0, m \text{\larger[1]{$)$}}^{2} \cap \mathbb{Z}^{2}$  for $i \in \{1,2\}$,  $a_{1}, a_{2} \in \mathbb{R}$, $t \in \mathbb{R}$, $x \in \mathbb{G}_{m}$ and $j \in \{1,\ldots,r\}$.  We recall as introduced in \eqref{fi}
\begin{align*}
f_{t}^{\scriptscriptstyle (i)}(x) := \mathbbm{1}_{\left\{\min\left(X_{x+e_{i}}, X_{\scriptscriptstyle x}\right) < t \leq \max\left(X_{x+e_{i}}, X_{\scriptscriptstyle x}\right)\right\}}
\end{align*}
Note that $0 \leq f_{t}^{\scriptscriptstyle (i)}(x) \leq 1$ \textit{a.s.}, let us denote the 4-dimensional Gaussian vector by $\mathbb{X}_{x_1, x_2} = (X_{x_1, x_2}, X_{x_1 + 1, x_2}, X_{x_2, x_1+1}, X_{x_2, x_1} )$ and $g_{t} : \mathbb{R}^{4} \to \mathbb{R}$, 
\begin{align*}
g_{t}\left(\mathbb{X}_{x_1, x_2}\right) & := a_{1}\mathbbm{1}_{\left\{\min\left(X_{x_1, x_2}, X_{\scriptscriptstyle x_1 +  1, \scriptscriptstyle x_2}\right) < t \leq \max\left(X_{x_1, x_2}, X_{\scriptscriptstyle x_1 +  1, \scriptscriptstyle x_2}\right)\right\}} \\
& \hspace{1cm} + a_{2}\mathbbm{1}_{\left\{\min\left(X_{x_2, x_1}, X_{\scriptscriptstyle x_2, x_1 + 1}\right) < t \leq \max\left(X_{x_2, x_1}, X_{\scriptscriptstyle x_2, x_1 + 1}\right)\right\}}
\end{align*}
Considering the sum over all elements of the window, we get that 
\begin{align*} 
\sum_{(x_1, x_2) \in \mathbb{G}_{m}} g_{t}\left(\mathbb{X}_{x_1, x_2}\right) - \mathbb{E}\left(g_{t}\left(\mathbb{X}_{x_1, x_2}\right)\right) = a_{1}\left(\mathcal{P}^{\scriptscriptstyle (1)}(t) - \mathbb{E}\left(\mathcal{P}^{\scriptscriptstyle (1)}(t)\right)\right) - a_{2}\left(\mathcal{P}^{\scriptscriptstyle (2)}(t) - \mathbb{E}\left(\mathcal{P}^{\scriptscriptstyle (2)}(t)\right)\right),
\end{align*}
Applying Theorem $2$ in \cite{arcones} we get that 
\begin{align*}
m^{-1} \sum_{(x_1, x_2) \in \mathbb{G}_{m}} f_{\scriptscriptstyle t}\left(\mathbb{X}_{x_1, x_2}\right) - \mathbb{E}\left(f_{\scriptscriptstyle t}\left(\mathbb{X}_{x_1, x_2}\right)\right) \xrightarrow[m \to \infty]{d} \mathcal{N}\left(0,\,\sigma^{2}\right),
\end{align*}
with 
\begin{align*}
\sigma^{2} := \mathbb{E}\left(\big[f_{t}(\mathbb{X}_{0}) - \mathbb{E}\left(f_{t}(\mathbb{X}_{0})\right)\big]^{2}\right) \\
& \hspace{-4cm}+ 2 \sum_{x_{1}, x_{2} = 1}^{\infty} \mathbb{E}\left(\big[f_{t}(\mathbb{X}_{0}) - \mathbb{E}\left(f_{t}(\mathbb{X}_{0})\right)\big]\big[f_{t}(\mathbb{X}_{x_1,x_2}) - \mathbb{E}\left(f_{t}(\mathbb{X}_{x_1, x_2})\right)\big]\right)
\end{align*}
\end{proof}
\begin{remark}
It is possible to generalize the result for $r$-thresholds by applying the same procedure.
\end{remark}
\paragraph{A test statistic} Following the idea  in \cite{bierme2019}  we propose a method to test the isotropy of the field $X$.
Let us consider the null hypothesis
\begin{center}
  $H_{0}$ \, :\; The field $X$ is isotropic,
\end{center}
and the following normality asymptotic result that is obtained by applying the delta method to Theorem (\textcolor{blue}{\ref{TCL}}).
\begin{Proposition}[Ratio of oriented perimeters]\label{deltamethodgeneralprop} 
Let  $t \in \mathbb{R}$, and $\check{\mathcal{P}}^{\scriptscriptstyle  (i) }(\cdot)$ be the scaled oriented perimeter over direction $i$, introduced in Definition \ref{defPerimetre} computed at level $t$. Then, it holds that,
\begin{equation*}
\frac{1}{m}\left(\cfrac{\mathcal{P}^{\scriptscriptstyle  (1) }_{m}(t)}{\mathcal{P}^{\scriptscriptstyle  (2)}_{m}(t)} - \cfrac{\mathbb{E}\left(\mathcal{P}^{\scriptscriptstyle  (1) }_{m}(t)\right)}{\mathbb{E}\left(\mathcal{P}^{\scriptscriptstyle  (2)}_{m}(t)\right)} \right)  \xrightarrow[m \to \infty]{d} \mathcal{N}\left(0,  \tilde{\sigma}^{2}(t) \right),
\end{equation*}
with $\tilde{\sigma}^{2}(t) \in (0, \infty)$.
% with $\tilde{\sigma}^{2}(t) = \frac{\tilde{\sigma}_{\scriptscriptstyle 1,1}^{2}(t)(\mu^{\star}_{2}(t))^{2} - 2\mu^{\star}_{2}(t)\mu^{\star}_{1}(t)\tilde{\sigma}_{\scriptscriptstyle 1,2}^{2}(t) +\tilde{\sigma}_{\scriptscriptstyle 2,2}^{2}(t)\mu^{\star}_{1}(t))^{2}}{\mu^{\star}_{2}(t)^4}$
Let $t \in \mathbb{R}$  and  $R_{m}(t) := \cfrac{\mathcal{P}^{\scriptscriptstyle  (1) }(t)}{\mathcal{P}^{\scriptscriptstyle  (2) }(t)}$. Then, under $H_{0}$, it holds that
\begin{equation}\label{ratiounderH0}
\frac{1}{m}\left(R_{m}(t) - 1 \right)  \xrightarrow[m \to \infty]{d, H_0} \mathcal{N}\left(0,  \tilde{\sigma}^{2}(t)  \right),
\end{equation} 
% $\tilde{\sigma}^{2}(t) = \frac{2\left(\tilde{\sigma}_{\scriptscriptstyle 1,1}^{2}(t) \, - \, \tilde{\sigma}_{\scriptscriptstyle 1,2}^{2}(t)\right)}{\mu^{\star}_{2}(t)^2}$
\end{Proposition}
Consider the consistent empirical estimator $\hat{\sigma}(t):= \frac{1}{m^2}\hat{\mathbb{V}\left(R_{m}(t)\right)}$ of $\tilde{\sigma}^{2}(t)$, where $\hat{\mathbb{V}\left(R_{m}(t)\right)}$ is the empirical variance of the considered ratio. Then, from \ref{deltamethodgeneralprop}, it holds that 
$$ \frac{1}{m\sqrt{\hat{\sigma}(t)}}\left(R_{m}(t) - 1 \right) = \frac{1}{\sqrt{\hat{\mathbb{V}\left(R_{m}(t)\right)}}}\left(R_{m}(t) - 1 \right)   \xrightarrow[m \to \infty]{d, H_{0}} \mathcal{N}\left(0, 1 \right).$$
Take a confidence level $\alpha \in (0,1)$ and set $q_{1-\frac{\alpha}{2}}$ such that $\mathbb{P}\left(|\mathcal{N}\left(0,1\right)| \leq  q_{1-\frac{\alpha}{2}} \right) = 1 - \frac{\alpha}{2}$. We define the symmetric accessible test $\hat{\phi}_{m}(t)$ with asymptotic level $\alpha$ as 
\begin{equation}
\label{phihat}
\hat{\phi}_{m}(t)= \mathbbm{1}_{\left\{ \frac{1}{\sqrt{\hat{\mathbb{V}\left(R_{m}(t)\right)}}}\left|\left(R_{m}(t) - 1 \right)\right| \geq q_{1- \frac{\alpha}{2}} \right\}}
\end{equation}
\begin{remark}(Power of the statistical test for the level $t = 0$) In Remark \ref{decreasingrho} we showed that the function $\rho \mapsto 2\mathbb{E}\left(\frac{\sqrt{1-\rho}}{\sqrt{1+\rho}}|N|\right) - 1 $ is strictly decreasing, thus, if $\rho(e_1) \neq \rho(e_2)$ thus $\frac{\mathbb{E}\left(\mathcal{P}^{1}(0)\right)}{\mathbb{E}\left(\mathcal{P}^{2}(0)\right)} \neq 1$. 
\end{remark}
\section{Numerical studies}
\subsection{Numerical studies of moments of the estimated perimeter}

Let $X$ be a stationary, isotropic centered Gaussian field with covariance function given by $\rho(x) = (1 -3||x||_{2}/3C + (||x||_{2}/2C)^{3})\mathbbm{1}_{||x||_{2} < C}, x \in \mathbb{R}^2$
Let ${Y(x; a, b, \theta): x \in \mathbb{R}^{2}}$ be a random field equal in distribution to  $\{X(Ax): x\in \mathbb{R}^{2}\}$, where 
\begin{equation}
A:= \begin{pmatrix} a & 0 \\ 0 & b\end{pmatrix} \begin{pmatrix} \cos(\theta) & \sin(\theta) \\ -\sin(\theta) & \cos(\theta)\end{pmatrix} 
\end{equation}
with $a, b \in \mathbb{R}^{2}$ and $\theta \in [0, \pi)$, the field $Y$ is also Gaussian with covariance function given by  $\rho_{X}(x) = (1 -3||Ax||_{2}/3C + (||Ax||_{2}/2C)^{3})\mathbbm{1}_{||Ax||_{2} < C},$, note that $a\neq b$ if and only if $Y$ is anisotropic. We chose to illustrate the computations of the statistics of the perimeter, using this covariance structure, and the value $\theta = 0$ and $a = 1$. 
\begin{figure}[H]
  \centering
    {\includegraphics[scale=0.3]{EsperanceUnedirection.pdf}}
    \hspace{0.2cm}
 \caption{In each panel, we represent the values of first moment of the oriented perimeter in direction $2$ computed for different values of $b$, the theoretical curve Equation \eqref{propFisrtmoment} is depicted using a blue color (blue) and the empirical values are in orange, the computations were done with a Montecarlo $= 100000$, for different values of $b = 1, 2.5, 5, 10$ and $C = 40$. }
\label{fig2}
\end{figure}

\begin{figure}[H]
  \centering
    {\includegraphics[scale=0.3]{EsperanceUnedirectionfonctionB.pdf}}
    \hspace{0.2cm}
 \caption{In each panel, we represent the values of first moment of the oriented perimeter in direction $2$ computed as a fonction of $b$ for a fixed level $t = 0$, the theoretical curve Equation \eqref{propFisrtmoment} is depicted using a blue color (blue) and the empirical values are in orange, the computations were done with a Montecarlo $= 100000$ and $C = 40$.}
\label{fig2}
\end{figure}

\begin{figure}[H]
  \centering
    {\includegraphics[scale=0.3]{ThreeConfNew.pdf}}
    \hspace{0.2cm}
 \caption{In each panel, we represent of the values of $\mathbb{E}\left(f^{1}(0,1)f^{2}(x)\right)$ for $x = (1,-1)$ computed for different values of $b$, the value of the theoretical curve Equation \eqref{thisthothree} in (blue)and the empirical values are in orange, the computations were done with a Montecarlo $= 100000$, for different values of $b = 1, 2.5, 5, 10$ and $C = 40$. }
\label{fig2}
\end{figure}

\begin{figure}[H]
  \centering
    {\includegraphics[scale=0.3]{fourConfNew.pdf}}
    \hspace{0.2cm}
 \caption{In each panel, we represent of the values of $\mathbb{E}\left(f^{1}(0,1)f^{2}(x)\right)$ for $x = (1, 1)$ computed for different values of $b$, the value of the theoretical curve Equation \eqref{thisthothree} in (blue)and the empirical values are in orange, the computations were done with a Montecarlo $= 100000$, for different values of $b = 1, 2.5, 5, 10$ and $C = 40$. }
\label{fig2}
\end{figure}

\begin{figure}[H]
  \centering
    {\includegraphics[scale=0.3]{Variance_40_30_10_1.pdf}}
    \hspace{0.2cm}
 \caption{We represent of the values of $\mathbb{V}\left(\mathcal{P}\right)(t)$, the value of the theoretical curve Equation \eqref{Thisthofour} in (blue) and the empirical values are in orange, the computations were done with a Montecarlo $= 100000$ to compute the theoretical value and Montecarlo $= 2000$ to compute the empirical one, for the values $b = 1$ and $C = 40$. }
\label{fig2}
\end{figure}

\begin{figure}[H]
  \centering
    {\includegraphics[scale=0.4]{Testanisotropy2.pdf}}
    \hspace{0.2cm} 
 \caption{We represent the values of $\hat{\Phi}_{m}(b)$ see Equation \eqref{phihat} in blue color computed for a Montecarlo $= 2000$, for different values of $b = 1, 1.2, 1.5, 2$ and $C = 40$.}
\label{fig2}
\end{figure}

\begin{figure}[H]
  \centering
    {\includegraphics[scale=0.4]{TestaniPhi0.pdf}}
    \hspace{0.2cm} 
 \caption{We represent the values of $\hat{\Phi}_{m}$ see Equation \eqref{phihat}, computed as a function of $b$ and a fixed value $t=0$ in blue color for a Montecarlo $= 2000$, and $C = 40$.}
\label{fig2}
\end{figure}

\begin{figure}[H]
  \centering
    {\includegraphics[scale=0.2]{TestaniRatiovsRHO.pdf}}
    {\includegraphics[scale=0.2]{TestaniRatiovsTVA.pdf}}
    \hspace{0.2cm} 
 \caption{We represent the values of the mean of the ratio between the oriented perimeter $\mathbb{E}\left(\frac{\mathcal{P}^{(1)}(0)}{\mathcal{P}^{(2)}(0)}\right)$ in blue and the ratio between of the empirical value for $\hat{\rho(e_i)}$ computed using the first moment of the total variation (\ref{totalvariation}) for different values of $b$, Montecarlo $= 500$, $t = 0$ and $C = 40$.}
\label{fig2}
\end{figure}

\begin{figure}[H]
  \centering
    {\includegraphics[scale=0.3]{imageU.pdf}}
    \hspace{3cm}
    {\includegraphics[scale=0.3]{levelset30000.pdf}}
    \hspace{0.2cm}
 \caption{ Image of level $30000$}
\label{fig2}
\end{figure}

\begin{figure}[H]
  \centering
    {\includegraphics[scale=0.25]{ConfidenceIntervalRDStD.pdf}}
    {\includegraphics[scale=0.25]{BoxplotRDStD.pdf}}
    \hspace{0.2cm}
 \caption{We represent the mean values of the ratio between the two oriented perimeters $\mathcal{P}^{1}, \mathcal{P}^{2}$ for real life data.}
\label{fig2}
\end{figure}

\begin{figure}[H]
  \centering
    {\includegraphics[scale=0.25]{ConfidenceIntervalRDAu.pdf}}
    {\includegraphics[scale=0.25]{BoxplotRDAu.pdf}}
    \hspace{0.2cm}
 \caption{We represent the mean values of the ratio between the two oriented perimeters $\mathcal{P}^{1}, \mathcal{P}^{2}$ for real life data.}
\label{fig2}
\end{figure}
\appendix 
\section{Computation of the covariance for the three cell configuration}
Let us consider the particular case where $x \in \left\{-e_1, e_1, -e_2, e_2 \right\}$, and consider the Gaussian vector $\left(X_{0}, X_{e_1}, X_{x}, X_{x+e_i} \right)$ and its associated transformation $\left(\Delta^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}, S^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}, S^{\scriptscriptstyle (i)}_{x} \right)$ of covariance matrix $\tilde{\Sigma}_{3}(x)$ again assuming that $|\rho(e_1)| < 1$, the Cholesky decomposition of $\tilde{\sigma}_{3}(x)$ allows us to get the expression of the Gaussian vector $\left(\Delta^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0},S^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0},X_{x} \right)$ in a new basis $\left(U,V,W,Z\right)$ with $\left(U, V, W\right) \sim \mathcal{N}\left(0,I_{3}\right)$. We get the following decomposition \\ 
$\left\{
\begin{array}{rlc}
\Delta^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0} & = & \sigma^{\scriptscriptstyle (-)}_{1}U,  \textcolor{white}{+  \pi^{\scriptscriptstyle(+, i)}_{V}(x) V +  \pi^{\scriptscriptstyle(+, i)}_{W}(x) W +  \pi^{\scriptscriptstyle(+, i)}_{Z}(x)Z, } \\
S^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0} & = & \textcolor{white}{\sigma^{\scriptscriptstyle (-)}_{1}U,} \sigma^{\scriptscriptstyle (+)}_{1}V, \textcolor{white}{+  \pi^{\scriptscriptstyle(+, i)}_{W}(x) W +  \pi^{\scriptscriptstyle(+, i)}_{Z}(x)Z, } \\
S^{\scriptscriptstyle (i)}_{\scriptscriptstyle  x} & = &  \pi^{\scriptscriptstyle(i)}_{U}(x)U +  \pi^{\scriptscriptstyle(i)}_{V}(x) V +  \pi^{\scriptscriptstyle(i)}_{W}(x) W, \textcolor{white}{ + \pi^{\scriptscriptstyle(+, i)}_{Z}(x)Z, } \\
\end{array}\right.$ \\
% with 
%  $\left\{
%  \begin{array}{rlc}
%  \pi_{U}(x) & = & \frac{1}{\sigma^{\scriptscriptstyle (-)}_{1}}\left(\rho(x-e_1) - \rho(x)\right) \\
%  \pi_{V}(x) & = &  \frac{1}{\sigma^{\scriptscriptstyle (+)}_{1}}\left(\rho(x-e_1) + \rho(x)\right)  \\
%  \pi_{W}(x)& = & \left(1 - \pi_{U}(x)^{2} - \pi_{V}(x)^{2}\right)^{\scriptscriptstyle (1/2)} \end{array}\right.$ 
Considering the first part of the previous equation, and applying Lemma \ref{lemmasommediff}
{\small
\begin{align*}
\mathbb{E}\left(\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle 0}, X_{e_1}) \leq t \leq \max(X_{\scriptscriptstyle 0}, X_{e_1})\right\}}\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle x}, X_{0}) \leq s \leq \max(X_{\scriptscriptstyle x}, X_{0}) \right\}}\right)\\
& \hspace{-6cm} = \mathbb{E}\left(\mathbbm{1}_{\left\{ \left|2t - S^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}\right| \leq \left|\Delta^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}\right|\right\}}\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle x}+ X_{e_1}, X_{0} + X_{e_1}) \leq s + X_{e_1}\leq \max(X_{\scriptscriptstyle x} + X_{e_1}, X_{0} + X_{e_1}) \right\}}\right). \\
& \hspace{-6cm} = \mathbb{E}\left(\mathbbm{1}_{\left\{ \left|2t - S^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}\right| \leq \left|\Delta^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}\right|\right\}}\mathbbm{1}_{\left\{\min(S^{(1)}_{x}, S^{(1)}_{0}) \leq s + \frac{S^{(1)}_{0} + \Delta^{(1)}_{0}}{2}\leq \max(S^{(1)}_{x}, S^{(1)}_{0}) \right\}}\right) \\
& \hspace{-6cm} = \mathbb{E}\left(\mathbbm{1}_{\left\{ \left|2t - S^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}\right| \leq \left|\Delta^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}\right|\right\}}\mathbbm{1}_{\left\{ \left|2s + \Delta^{(1)}_{0} -  S^{\scriptscriptstyle (1)}_{x}  \right| \leq \left|S^{\scriptscriptstyle (1)}_{x} - S^{\scriptscriptstyle (1)}_{0}\right|\right\}}\right). 
\end{align*}}
besides, 
\begin{align*}
\mathbbm{1}_{\left\{ \left|2s + \Delta^{(1)}_{0} -  S^{\scriptscriptstyle (1)}_{x}  \right| \leq \left|S^{\scriptscriptstyle (1)}_{x} - S^{\scriptscriptstyle (1)}_{0}\right|\right\}} &= \mathbbm{1}_{\left\{  S^{\scriptscriptstyle (1)}_{0} \leq 2s + \Delta^{(1)}_{0}\right\}} \mathbbm{1}_{\left\{ S_{x} \geq  \max\left(s+ \frac{1}{2}S^{\scriptscriptstyle (1)}_{0} + \frac{1}{2}\Delta^{\scriptscriptstyle (1)}_{0}, S^{\scriptscriptstyle (1)}_{0}\right) \right\}} \\
& + \mathbbm{1}_{\left\{  2s + \Delta^{(1)}_{0} \leq S^{\scriptscriptstyle (1)}_{0} \right\}} \mathbbm{1}_{\left\{ S_{x} \leq  \min\left(s+ \frac{1}{2}S^{\scriptscriptstyle (1)}_{0} + \frac{1}{2}\Delta^{\scriptscriptstyle (1)}_{0}, S^{\scriptscriptstyle (1)}_{0}\right) \right\}} \\
&= \mathbbm{1}_{\left\{  S^{\scriptscriptstyle (1)}_{0} \leq 2s + \Delta^{(1)}_{0}\right\}} \mathbbm{1}_{\left\{ S_{x} \geq s+ \frac{1}{2}S^{\scriptscriptstyle (1)}_{0} + \frac{1}{2}\Delta^{\scriptscriptstyle (1)}_{0}\right\}} + \mathbbm{1}_{\left\{  2s + \Delta^{(1)}_{0} \leq S^{\scriptscriptstyle (1)}_{0} \right\}} \mathbbm{1}_{\left\{ S_{x} \leq  S^{\scriptscriptstyle (1)}_{0} \right\}} \\
\end{align*} 
% \begin{align*}
% \mathbbm{1}_{\left\{ S^{\scriptscriptstyle (1)}_{x} - S^{\scriptscriptstyle (1)}_{0} \geq 0 \right\}} \mathbbm{1}_{\left\{  - S^{\scriptscriptstyle (1)}_{x} + S^{\scriptscriptstyle (1)}_{0} \leq 2s + \Delta^{(1)}_{0} -  S^{\scriptscriptstyle (1)}_{x}  \leq S^{\scriptscriptstyle (1)}_{x} - S^{\scriptscriptstyle (1)}_{0}\right\}} & =\\
% & \hspace{-4.5cm} \mathbbm{1}_{\left\{ S^{\scriptscriptstyle (1)}_{x}  \geq S^{\scriptscriptstyle (1)}_{0}\right\}} \mathbbm{1}_{\left\{  S^{\scriptscriptstyle (1)}_{0} \leq 2s + \Delta^{(1)}_{0}\right\}}\mathbbm{1}_{\left\{ 2s + \Delta^{(1)}_{0}  + S^{\scriptscriptstyle (1)}_{0} \leq 2S^{\scriptscriptstyle (1)}_{x} \right\}}\\
% & \hspace{-5cm} = \mathbbm{1}_{\left\{  S^{\scriptscriptstyle (1)}_{0} \leq 2s + \Delta^{(1)}_{0}\right\}} \mathbbm{1}_{\left\{\pi^{\scriptscriptstyle(i)}_{W}(x) W \geq  - \pi^{\scriptscriptstyle(i)}_{U}(x)U + \left(\sigma^{(+)}_{1} - \pi^{\scriptscriptstyle(i)}_{V}(x)\right)V\right\}}\\
% & \hspace{1cm} \mathbbm{1}_{\left\{ 2s + \left(\sigma^{(-)}_{1} - 2\pi^{\scriptscriptstyle(i)}_{U}(x)\right)U  +  \left(\sigma^{(+)}_{1} - 2\pi^{\scriptscriptstyle(i)}_{V}(x)\right) \leq 2\pi^{\scriptscriptstyle(i)}_{W}(x) W \right\}} \\
% & \hspace{-5cm} = \mathbbm{1}_{\left\{  S^{\scriptscriptstyle (1)}_{0} \leq 2s + \Delta^{(1)}_{0}\right\}} \\
% & \hspace{-4cm} \mathbbm{1}_{\left\{\pi^{\scriptscriptstyle(i)}_{W}(x) W  \geq  \max(- \pi^{\scriptscriptstyle(i)}_{U}(x)U + \left(\sigma^{(+)}_{1} - \pi^{\scriptscriptstyle(i)}_{V}(x)\right)V, s + \left(\frac{1}{2}\sigma^{(-)}_{1}- \pi^{\scriptscriptstyle(i)}_{U}(x)\right)U  +  \left(\frac{1}{2}\sigma^{(+)}_{1}- \pi^{\scriptscriptstyle(i)}_{V}(x)\right) \right\}}\\
% \end{align*}

% \begin{align*}
% \mathbbm{1}_{\left\{ S^{\scriptscriptstyle (1)}_{x} - S^{\scriptscriptstyle (1)}_{0} \leq 0 \right\}} \mathbbm{1}_{\left\{  S^{\scriptscriptstyle (1)}_{x} - S^{\scriptscriptstyle (1)}_{0} \leq 2s + \Delta^{(1)}_{0} -  S^{\scriptscriptstyle (1)}_{x}  \leq -S^{\scriptscriptstyle (1)}_{x} + S^{\scriptscriptstyle (1)}_{0}\right\}} & =\\
% & \hspace{-4.5cm} \mathbbm{1}_{\left\{ S^{\scriptscriptstyle (1)}_{x}  \leq S^{\scriptscriptstyle (1)}_{0}\right\}} \mathbbm{1}_{\left\{  S^{\scriptscriptstyle (1)}_{0} \geq 2s + \Delta^{(1)}_{0}\right\}}\mathbbm{1}_{\left\{ 2s + \Delta^{(1)}_{0}  + S^{\scriptscriptstyle (1)}_{0} \geq 2S^{\scriptscriptstyle (1)}_{x} \right\}}\\
% & \hspace{-5cm} = \mathbbm{1}_{\left\{  S^{\scriptscriptstyle (1)}_{0} \geq 2s + \Delta^{(1)}_{0}\right\}} \mathbbm{1}_{\left\{\pi^{\scriptscriptstyle(i)}_{W}(x) W \leq  - \pi^{\scriptscriptstyle(i)}_{U}(x)U + \left(\sigma^{(+)}_{1} - \pi^{\scriptscriptstyle(i)}_{V}(x)\right)V\right\}}\\
% & \hspace{-1.7cm} \mathbbm{1}_{\left\{ 2s + \left(\sigma^{(-)}_{1} - 2\pi^{\scriptscriptstyle(i)}_{U}(x)\right)U  +  \left(\sigma^{(+)}_{1} - 2\pi^{\scriptscriptstyle(i)}_{V}(x)\right) \geq 2\pi^{\scriptscriptstyle(i)}_{W}(x) W \right\}} \\
% & \hspace{-5cm} = \mathbbm{1}_{\left\{  S^{\scriptscriptstyle (1)}_{0} \geq 2s + \Delta^{(1)}_{0}\right\}} \\
% & \hspace{-4cm} \mathbbm{1}_{\left\{\pi^{\scriptscriptstyle(i)}_{W}(x) W \leq  \min(- \pi^{\scriptscriptstyle(i)}_{U}(x)U + \left(\sigma^{(+)}_{1} - \pi^{\scriptscriptstyle(i)}_{V}(x)\right)V, s + \left(\frac{1}{2}\sigma^{(-)}_{1}- \pi^{\scriptscriptstyle(i)}_{U}(x)\right)U  +  \left(\frac{1}{2}\sigma^{(+)}_{1}- \pi^{\scriptscriptstyle(i)}_{V}(x)\right) \right\}}\\
% \end{align*}
Thus, putting everything together, we get 
{\small 
\begin{align*}
\mathbb{E}\left(\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle 0}, X_{e_1}) \leq t \leq \max(X_{\scriptscriptstyle 0}, X_{e_1})\right\}}\mathbbm{1}_{\left\{\min(X_{x}, X_{0}) \leq s \leq \max(X_{x}, X_{0}) \right\}}\right) = \\
& \hspace{-10cm} \mathbb{E}\left(\mathbbm{1}_{\left\{|2t - \sigma^{(+)}_{1}V \leq  \sigma^{(-)}_{1}U|\right\}}\mathbbm{1}_{\left\{ \sigma^{(+)}_{1}V  \leq 2s + \sigma^{(-)}_{1}U\right\}} \left(1 - \Phi_{\pi^{\scriptscriptstyle(i)}_{W}(x) }\left(s + \left(\frac{1}{2}\sigma^{(+)}_{1}- \pi^{\scriptscriptstyle(i)}_{U}(x)\right)U  +  \left(\frac{1}{2}\sigma^{(+)}_{1}- \pi^{\scriptscriptstyle(i)}_{V}(x)\right)V\right)\right)\right)\\
& \hspace{-10cm}+ \mathbb{E}\left(\mathbbm{1}_{\left\{|2t - \sigma^{(+)}_{1}V \leq  \sigma^{(-)}_{1}U|\right\}}\mathbbm{1}_{\left\{ \sigma^{(+)}_{1}V  \geq 2s + \sigma^{(-)}_{1}U\right\}}\Phi_{\pi^{\scriptscriptstyle(i)}_{W}(x) }\left(- \pi^{\scriptscriptstyle(i)}_{U}(x)U + \left(\sigma^{(+)}_{1} - \pi^{\scriptscriptstyle(i)}_{V}(x)\right)V\right) \right)\\\
\end{align*}
}
The second configuration would be to have
{\small
\begin{align*}
\mathbb{E}\left(\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle 0}, X_{e_1}) \leq t \leq \max(X_{\scriptscriptstyle 0}, X_{e_1})\right\}}\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle x}, X_{e_1}) \leq s \leq \max(X_{\scriptscriptstyle x}, X_{e_1}) \right\}}\right)\\
& \hspace{-6cm} = \mathbb{E}\left(\mathbbm{1}_{\left\{ \left|2t - S^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}\right| \leq \left|\Delta^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}\right|\right\}}\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle x}+ X_{0}, X_{e_1} + X_{0}) \leq s + X_{0}\leq \max(X_{\scriptscriptstyle x} + X_{0}, X_{0} + X_{e_1}) \right\}}\right). \\
& \hspace{-6cm} = \mathbb{E}\left(\mathbbm{1}_{\left\{ \left|2t - S^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}\right| \leq \left|\Delta^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}\right|\right\}}\mathbbm{1}_{\left\{\min(S^{(1)}_{x}, S^{(1)}_{0}) \leq s + \frac{S^{(1)}_{0} - \Delta^{(1)}_{0}}{2}\leq \max(S^{(1)}_{x}, S^{(1)}_{0}) \right\}}\right) \\
& \hspace{-6cm} = \mathbb{E}\left(\mathbbm{1}_{\left\{ \left|2t - S^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}\right| \leq \left|\Delta^{\scriptscriptstyle (1)}_{\scriptscriptstyle 0}\right|\right\}}\mathbbm{1}_{\left\{ \left|2s - \Delta^{(1)}_{0} -  S^{\scriptscriptstyle (1)}_{x}  \right| \leq \left|S^{\scriptscriptstyle (1)}_{x} - S^{\scriptscriptstyle (1)}_{0}\right|\right\}}\right). 
\end{align*}}
besides, 
besides, 
\begin{align*}
\mathbbm{1}_{\left\{ \left|2s - \Delta^{(1)}_{0} -  S^{\scriptscriptstyle (1)}_{x}  \right| \leq \left|S^{\scriptscriptstyle (1)}_{x} - S^{\scriptscriptstyle (1)}_{0}\right|\right\}} 
&= \mathbbm{1}_{\left\{  S^{\scriptscriptstyle (1)}_{0} \leq 2s - \Delta^{(1)}_{0}\right\}} \mathbbm{1}_{\left\{ S_{x} \geq s+ \frac{1}{2}S^{\scriptscriptstyle (1)}_{0} - \frac{1}{2}\Delta^{\scriptscriptstyle (1)}_{0}\right\}} + \mathbbm{1}_{\left\{  2s - \Delta^{(1)}_{0} \leq S^{\scriptscriptstyle (1)}_{0} \right\}} \mathbbm{1}_{\left\{ S_{x} \leq  S^{\scriptscriptstyle (1)}_{0} \right\}} \\
\end{align*} 
Thus, putting everything together, we get 
{\small 
\begin{align*}
\mathbb{E}\left(\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle 0}, X_{e_1}) \leq t \leq \max(X_{\scriptscriptstyle 0}, X_{e_1})\right\}}\mathbbm{1}_{\left\{\min(X_{x}, X_{0}) \leq s \leq \max(X_{x}, X_{0}) \right\}}\right) = \\
& \hspace{-10cm}\mathbb{E}\left(\mathbbm{1}_{\left\{|2t - \sigma^{(+)}_{1}V \leq  \sigma^{(-)}_{1}U|\right\}}\mathbbm{1}_{\left\{ \sigma^{(+)}_{1}V  \leq 2s - \sigma^{(-)}_{1}U\right\}}\right. \\
& \hspace{-10cm}\left. \left(1 - \Phi_{\pi^{\scriptscriptstyle(i)}_{W}(x) }\left(\max\left(- \pi^{\scriptscriptstyle(i)}_{U}(x)U + \left(\sigma^{(+)}_{1} - \pi^{\scriptscriptstyle(i)}_{V}(x)\right)V, s - \left(\frac{1}{2}\sigma^{(-)}_{1}+ \pi^{\scriptscriptstyle(i)}_{U}(x)\right)U  +  \left(\frac{1}{2}\sigma^{(+)}_{1}- \pi^{\scriptscriptstyle(i)}_{V}(x)\right)V\right)\right) \right)\right)\\
& \hspace{-10cm}+ \mathbb{E}\left(\mathbbm{1}_{\left\{|2t - \sigma^{(+)}_{1}V \leq  \sigma^{(-)}_{1}U|\right\}}\mathbbm{1}_{\left\{ \sigma^{(+)}_{1}V  \geq 2s + \sigma^{(-)}_{1}U\right\}}\right. \\
& \hspace{-10cm}\left. \Phi_{\pi^{\scriptscriptstyle(i)}_{W}(x) }\left(\min\left(- \pi^{\scriptscriptstyle(i)}_{U}(x)U + \left(\sigma^{(+)}_{1} - \pi^{\scriptscriptstyle(i)}_{V}(x)\right)V, s - \left(\frac{1}{2}\sigma^{(-)}_{1}+ \pi^{\scriptscriptstyle(i)}_{U}(x)\right)U  +  \left(\frac{1}{2}\sigma^{(+)}_{1}- \pi^{\scriptscriptstyle(i)}_{V}(x)\right)V\right)\right) \right)\\\
\end{align*}
}

Regarding the three cell configuration, one would have to reconsider the expression from the beginning and integrate the third variable $W$ that has a non vanishing variance in the independent regime \textit{i.e.} $\rho(e_1) = \rho(2e_1)=0$. Thus, those computations would give us the following expression,
{\tiny
\begin{align}
\label{IIDCase}
\mathbb{E}\left(\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle 0}, X_{e_1}) \leq t \leq \max(X_{\scriptscriptstyle 0}, X_{e_1})\right\}}\mathbbm{1}_{\left\{\min(X_{e_1}, X_{e_2}) \leq s \leq \max(X_{e_1}, X_{2e_1}) \right\}}\right) \nonumber \\
& \hspace{-8.7cm} = \mathbb{E}\left(\mathbbm{1}_{\left\{|2t - \sqrt{2(1+\rho(e_1))}V \leq \sqrt{2(1-\rho(e_1))}U|\right\}}\mathbbm{1}_{\left\{\sqrt{2(1-\rho(e_1))}U + \sqrt{2(1+\rho(e_1))}V \leq s\right\}}\right. \nonumber \\
& \hspace{-8cm} \left. \left(1 - \Phi_{\alpha^{(1)}(e_1)}\left(\max\left(\frac{1 + \rho(2e_1) + 2\rho(e_1)}{\sqrt{2(1-\rho(e_1))}}U + \frac{1-\rho(2e_1)}{\sqrt{2(1+\rho(e_1))}}V, s - \frac{\rho(e_1) - \rho(2e_1)}{\sqrt{2(1+\rho(e_1))}} U - \frac{\rho(e_1) + \rho(2e_1)}{\sqrt{2(1+\rho(e_1))}}V \right) \right)\right)\right) \nonumber\\
& \hspace{-8cm} + \mathbb{E}\left(\mathbbm{1}_{\left\{|2t - \sqrt{2(1+\rho(e_1))}V \leq \sqrt{2(1-\rho(e_1))}U|\right\}}\mathbbm{1}_{\left\{\sqrt{2(1-\rho(e_1))}U+ \sqrt{2(1+\rho(e_1))}V \geq s\right\}}\right. \nonumber\\
& \hspace{-8cm} \left. \left(\Phi_{\alpha^{(1)}(e_1)}\left(\min\left(\frac{1 + \rho(2e_1) + 2\rho(e_1)}{\sqrt{2(1-\rho(e_1))}}U + \frac{1-\rho(2e_1)}{\sqrt{2(1+\rho(e_1))}}V, s - \frac{\rho(e_1) - \rho(2e_1)}{\sqrt{2(1+\rho(e_1))}} U - \frac{\rho(e_1) + \rho(2e_1)}{\sqrt{2(1+\rho(e_1))}}V \right) \right)\right)\right),
\end{align}}then if the Gaussian field $\left(X(x)\right)_{x \in \mathbb{G}_{m}}$ is such that all the variables are independent, then $\rho(2e_1) = \rho(e_1) = 0$, thus Equation \eqref{IIDCase} becomes 
{\small
\begin{align*}
\mathbb{E}\left(\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle 0}, X_{e_1}) \leq t \leq \max(X_{\scriptscriptstyle 0}, X_{e_1})\right\}}\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle x}, X_{x+e_i}) \leq s \leq \max(X_{e_1}, X_{2e_1}) \right\}}\right) & \\
& \hspace{-9cm} = \mathbb{E}\left(\mathbbm{1}_{\left\{ |t - \frac{\sqrt{2}V}{2}| \leq |\frac{\sqrt{2}U}{2}|\right\}}\mathbbm{1}_{\left\{\frac{1}{\sqrt{2}} U + \frac{1}{\sqrt{2}}V \leq s\right\}} \left(1 - \Phi\left(s\right)\right)\right) \\
& \hspace{-9cm} + \mathbb{E}\left(\mathbbm{1}_{\left\{ |t - \frac{\sqrt{2}V}{2}| \leq |\frac{\sqrt{2}U}{2}|\right\}}\mathbbm{1}_{\left\{ s \leq \frac{1}{\sqrt{2}}U + \frac{1}{\sqrt{2}}V \right\}}\Phi\left(s\right)\right) \\
& \hspace{-9cm} = (1-\Phi(s))\mathbb{E}\left(\mathbbm{1}_{\left\{\frac{V-U}{\sqrt{2}} \leq t \right\}}\mathbbm{1}_{\left\{\frac{V+U}{\sqrt{2}} \geq t \right\}}\mathbbm{1}_{\left\{U \geq 0\right\}}\left(\mathbbm{1}_{\left\{\frac{V+U}{\sqrt{2}} \leq s \right\}} + \mathbbm{1}_{\left\{\frac{V-U}{\sqrt{2}} \leq s \right\}}\right) \right)\\
& \hspace{-9cm} + \Phi(s)\mathbb{E}\left(\mathbbm{1}_{\left\{\frac{V-U}{\sqrt{2}} \leq t \right\}}\mathbbm{1}_{\left\{\frac{V+U}{\sqrt{2}} \geq t \right\}}\mathbbm{1}_{\left\{U \geq 0\right\}}\left(\mathbbm{1}_{\left\{\frac{V+U}{\sqrt{2}} \geq s \right\}} + \mathbbm{1}_{\left\{\frac{V-U}{\sqrt{2}} \geq s \right\}}\right) \right)
\end{align*}}
Applying the same strategy we adopted in \ref{iidCase} and considering the change variable $Z = \frac{V-U}{\sqrt{2}}$ and $Y = \frac{U+V}{\sqrt{2}}$ we get 
{\small
\begin{align*}
\mathbb{E}\left(\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle 0}, X_{e_1}) \leq t \leq \max(X_{\scriptscriptstyle 0}, X_{e_1})\right\}}\mathbbm{1}_{\left\{\min(X_{\scriptscriptstyle x}, X_{x+e_i}) \leq s \leq \max(X_{e_1}, X_{2e_1}) \right\}}\right) & \\
& \hspace{-9cm} = (1-\Phi(s))\mathbb{E}\left(\mathbbm{1}_{\left\{Z \leq t \right\}}\mathbbm{1}_{\left\{Y \geq t \right\}}\left(\mathbbm{1}_{\left\{Y \leq s \right\}} + \mathbbm{1}_{\left\{Z \leq s \right\}}\right) \right)\\
& \hspace{-9cm} + \Phi(s)\mathbb{E}\left(\mathbbm{1}_{\left\{Z \leq t \right\}}\mathbbm{1}_{\left\{Y \geq t \right\}}\left(\mathbbm{1}_{\left\{Y \geq s \right\}} + \mathbbm{1}_{\left\{Z \geq s \right\}}\right) \right) \\
& \hspace{-9cm}= (1-\Phi(s))\Phi(t)\left(\Phi(s) - \Phi(t)\right) + (1-\Phi(s))\left(1-\Phi(t)\right)\Phi(t)   + \Phi(s)\Phi(t)\left(1-\Phi(s)\right)\\
& \hspace{-9cm}= \Phi(t)\left(1-\Phi(s)\right)\left(1+ 2\Phi(s) - 2\Phi(t)\right).
\end{align*}}
This result was previously proved in \cite{Psymetrie} (Proposition \textcolor{blue}{3.2}).
\bibliographystyle{plain}
\bibliography{biblioLKC}
\end{document}